### 2022-02-01 outputs/paper/paper.Rmd
The first commit. Trial run.

---
title: "Income Inequality is revealed in Torontonian Apartment Maintenance"
subtitle: "My subtitle if needed"
author: 
  - Olaedo Okpareke
  - Another author
thanks: "Code and data are available at: github repo. What addres do i put?."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Housing maintenance beneifts residents as it makes sure all ammenities and appliances function properly, maintains structural integrity, and adds value to homes. Previous research has found that private apartments tend to be better maintained than their social housing counterparts, and that housing in richer areas have better ammenities than those in poorer areas. We use data from Open data Toronto to further analyze this by breaking down apartment scores and other attributes by Ward and apartment type, using R (R Core Team 2021) and other packages. We find that apartments in higher income areas and privately owned apartments tend to have better maintenance scores than those in lower income wards and public housing respectively. This data shows the inequality in standards for housing in Toronto, as well as the need for better government accountability and housing justice. Improving the gap in aparment quality helps improve the life of poor and marginalized people."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---
BOXPLOTS TALK ABOYT MEDIANS< NOT MEANS. PUT THE TABLE FOR MEAN
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
install.packages("opendatatoronto")
library(opendatatoronto)
```

# 1. Introduction
Toronto is a large Canadian city divided into 25 wards. In order to make sure building maintenance standards are upheld and to allow apartment residents to look up safety scores of their apartments, Rent Safe TO has provides a dataset which, according to a set criteria not revealed to public, rates apartment attributes according to a 5 point system, and then uses a formula to assign them an overall score over 100. In this report, we are interested in how apartment maintenance scores have changed over the years, and how attributes such as relative wealth of the ward and the type of apartment affects the total score given to an apartment.

We obtained the dataset from the City of Toronto Open Data Portal [the address of this book](https://open.toronto.ca/dataset/apartment-building-evaluation/) with `opendatatoronto` (Gelfand 2020), and cleaned and analyzed the dataset using R (R Core Team 2021),tidyverse (Wickham et al. 2019) and dplyr (Wickham et al. 2021) packages. Figures and tables were created with ggplot2 (Wickham 2016) and kableextra (Zhu 2021). We created tables that first showed the data and important summary statistics, then went into showing summaries for area and apartment types, as well as box plots, linear model graphs and percentage bar charts that highlighted inequality in housing scores.

Not all apartments are equal, as this report shows that although general housing maintenance has increased in recent years (OECD 2020), there is still a large difference in quality of apartments around the city. Social housing, both operated by the TCHC and by other institutions, have lower average maintenance scores and lower median security scores than their private counterparts. Apartments in some wealthier wards also have a higher average maintenance score than those in poorer wards. According to (OECD 2020) and (Desroches and Poland, 2021), these disparities are due to a decline in public investment.

These scores are widely used to rate apartments ((Carter, 2004: 12)), however biased scoring and issues in measurement could create inaccurate property pricing, or otherwise give false impressions on the apartment block, and thus the type of people living there(Konadu-Agyemang et al. 1994, pg 146). Low scores for public housing also exacerbate the already low view of people who need them (Konadu-Agyemang et al. 1994, pg 146). This data and the statistics shows that vulnerable populations such as low income and minorities suffer systemic marginalization in facets of society, even in the housing sector as well maintained, secure housing is still unattainable for the city' most vulnerable, questioning the extent of the achievements of social housing. There are also concerns on gentrification, bias in grading, and ...

# 2 Data
The data for this report was gotten from the Apartment Building Evaluation (Rent Safe TO 2022) from the Toronto Open Data Portal (city of Toronto 2022) using the R package opendatatoronto (Gelfand 2020). This data is compiled by RentSafeTO and was last refreshed on January 22nd, 2022. Rent Safe TO collects data on apartments in the GTA that are 3 or more stories or had 10 more units, and response is mandatory for qualifying apartments, i.e the population. The raw data includes 9758 apartment blocks from the Greater Toronto Area and their categorical characteristics (40 variables in total) such as location and Type of property, as well as numeric characteristics such as Security,ammenities and state of common areas. It also includes an overall score over 100 for each apartment block.

For data cleaning, we removed all rows with missing values as they indicated that the apartment did not have a certain room, ammenity, or characteristic, but we wanted an "all things being equal" analysis, the assumption that attributes wise, they were alike. That is, we wanted to compare apartments that a had poor/ lacking characteristic, not no characteristic at all. It was also helpful to see the full picture so we could see if there was bias in grading instead of it not being graded at all. The missing values seemed to be randomly dispersed about the dataset so removing them would not cause much deviation.

We included a variable if there was anything important that made it stand out for our analysis.  We omitted most variables in the original set as their direct effects were not important to the analysis, and the total score variable encompassed all of their individual scores. we first removed _id and RSN as they were merely ID variables. We also excluded the ward number variable as the ward names variable were easier to identify. We included both year built as we wanted to see the change in housing maintenance between older and newer apartments. We included score as it was the main determinant of maintenance standards, as well as security to observe the the differences in security between different groups of apartments, and we wanted to see its direct effect. We then included property type, which states whether an apartment was privately owned, owned by the Toronto Housing Corporation(THCH), or owned by some other non-private institution as our analysis focuses on social/affordable housing. The final dataframe included 487 observations with 5 variables.

```{r}
apartment_data = list_package_resources("4ef82789-e038-44ef-a478-a8f3590c3eb1") %>%
  filter(name=="Apartment Building Evaluation") %>%
  get_resource()

apartment_data = na_if(apartment_data, 'N/A')

rent_data <- apartment_data[complete.cases(apartment_data), ]

# Try  to filter by year evaluated and see if it gets what I want
rent_clean<- rent_data %>% 
  dplyr::select(Score = SCORE,
                Ward = WARDNAME,
                Year_built = YEAR_BUILT,
                Type= PROPERTY_TYPE,
                Security = SECURITY
  )
```

```{r}
rent_clean %>%
  type_convert(
  rent_clean,
  col_types = cols(
    Score = col_integer(),
    Ward = col_character(),
    Year_built = col_double(),
    Type = col_character(),
    Security = col_double()
  ),
  na = c("", "N/A"),
  guess_integer = FALSE
)
```
```{r}
rent_clean$Score = strtoi(rent_clean$Score)
rent_clean$Year_built = as.double(rent_clean$Year_built)
rent_clean$Security = as.double(rent_clean$Security)
```

We are interested in the relationships between property Type and Ward on Score and security, as well as the trend of Score over time. It shows a measure of income inequality in housing, and has possible implications for public policy. A preview of the data, (Table \@ref(tab:1)) using kableexrra(Zhu 2021) is shown below.
```{r 1}
x = head(rent_clean) %>%
  knitr::kable(col.names = c("Score", "Ward", "Year built","Type", "Security"), caption = "Toronto Apartment Evaluation Data")
x
```

(Table \@ref(tab:2)) shows the the summary statistics for the buildings. We can see from the table that the average building has good standards and security, but is a bit old. The average apartment has a better score than the median apartment, indicating that there are some apartments that are far above the average.However the mean security is lower the median, indicating some very low security apartments. Policy is needed to improve security for these apartments, and the disrepancies that some scores are far higher than most are put into question.

```{r 2}
rent_clean %>% 
  summarise(mean=mean(rent_clean$Score),median=median(rent_clean$Score), max = max(rent_clean$Score),min=min(rent_clean$Score))
```

(Figure @\ref(fig: 1)), using ggplot2(Wickham 2016), shows the average maintenance scores by ward. Cross referencing the income and demographics of each ward using (City of Toronto 2022) , we see that the higher income wards such as Toronto Saint Paul's and Toronto-Danforth (155,470 and 101,323 average yearly income respectively 72% white and 67% white respectively as of 2016) tend to have both higher means and medians than lower income wards, such as Davenport and Humber River Black Creek (80,807 and 65,458 average yearly income as of 2016). An exception seems to be Scarborough North which has a low yearly average income and high racial minority population but a high average apartment maintenance score. However note that the median is very small, implies that the sample in this population may not be representative of the true state of apartments in this ward. Although there are several outliers Scores for each ward, in general apartment housing is better in richer wards than in poorer wards. High income Wards also tend to have fewer racial minorities, with Toronto saint Paul and Toronto-Danforth having a 72% and 67% white population respectively.
```{r 1, fig.cap = "Boxplot of mean and median apartment scores by ward", echo = TRUE}
rent_clean %>%
  ggplot(aes(x = Ward,y = Score,group = Ward)) + 
  theme_minimal() +
  geom_boxplot() +
  coord_flip()+
  labs(title = "Wealthier wards tend to have better apartment maintenance than poorer",
       x = "Ward",
       y = "Apartment Score")
```
(Figure @\ref(fig:2)) shows the mean and median apartment scores by type.  The results show that private buildings have the best average maintenance quality, and those from the Toronto housing Corporation have the lowest average. The spread of quality seems to be equal for all three housing types, so even accounting for variation private housing is consistently better than their public counterparts. This shows a disparity in priority for apartment maintenance and upkeep as well as general housing inequality, and shows that government are not on top of social housing affairs, as those in social housing owned by the TCHC also deserve high average maintenance taking care of their social housing. However it should be noted that (non THCH owned) social housing maintenance is not very far behind private housing, showing a just upkeep of housing. However according to (), people in social housing often have to take maintenance and upkeep into their own hands. 
```{r 2,fig.cap = "Boxplot of mean and median apartment scores by Apartment Type", echo = TRUE }
rent_clean %>%
  ggplot(aes(x = Type,y = Score)) + 
  theme_minimal() +
  geom_boxplot() +
  labs(title = "Private Housing has better scores than social housing of thch and not thch",
       x = "Apartment Type",
       y = "Apartment Score")
```
(Table \@ref(tab:3))  shows the differences in security between each housing type. We see that overall the average maintenance scores are good for all types of apartment. However the median shows that both THCH are lower than that of the private buildings, and their average scores were driven up by  a few very good apartments. This means that we can expect that a large number of social housing (which houses mainly low income populations) do not have the best security, which is counter-intuitive as poorer areas where this housing is often located are often in higher need of security due to correlation with crime rates. (cite poor housing live in crime areas). This could be alleviated with multi-income neighbourhoods, which places many different types of housing that can accomodate different income groups, but NIMBY-ism has slowed the progress of these neighbourhoods.
```{r}
rent_clean %>%
  group_by(Type) %>%# We want to know the occupancy by month
  summarize(Mean = mean(Security),Median = median(Security)) %>%
  knitr::kable()
```

(Figure @\ref(fig:3)) shows the general trend of apartment maintenance scores for the 3 types according to the year they were built. Even though for all types newer buildings have better maintenance Scores than their older counterparts for all three types of apartment, private housing still gets consistently higher scores than the other two types for almost all the years. Most private apartments were built between the 1960s to 1970s while social housing was built between the 1980s to 2000s, so even the newer age of the social housing buildings is not enough to protect against lower rates of maintenance. This shows either the persistence in lack of government and public maintenance or persistent bias in scoring, which all tie into income inequality. It should be noted that the trend for private apartments may be heavily influenced by the points at the far right end, and that many private apartments also have low scores. We also removed an outlier, 1910 as it affected the accuracy of the trend for private apartments. 
```{r 3,fig.cap = "Scatterplots and regression line of trend of apartment scores by type over 1950 - 2021", echo = FALSE }
rent_clean %>%
  filter(Year_built >=1950) %>% # The only value before 1950 (1910, is an outlier)
  ggplot(aes(x = Year_built, y = Score, color = Type)) +
  geom_point() +
  geom_smooth(method = 'lm', se = F)+
  geom_abline()+
  theme_minimal() +
  labs(title = "Privately owned apartment scores have ongiong advantage over social counterparts",
       x = "Year",
       y = "Score",
       color = "Type") +
  scale_color_brewer(palette = "Dark2")
```

(Figure @\ref(fig:4)) shows that the wealthier wards have a larger percentage of their apartmentd scoring above 70 compared to lower income wards. Some of the poorer wards such as Humber River Black creek or Davenport do not have very highly rated apartments at all. The racial minority population in these two wards as of 2016 was 74% and 43% respectively (City of Toronto 2022).In 2016 the income for these wards was 65,548 and 80,807 dollars respectively while in the richer wards like Toronto Saint Paul and Etobicoke Centre it was 155,470 and 128,448 (City of Toronto 2022). Many other wards follow the same trend, with the wards with higher income having larger percentages of their apartments being above 70, with some lower income wards even having some percentage of their apartments have a score less than 50. This is another way to see the inequality in housing standards, and communicates that housing maintenance disproportionally caters to higher income areas, possibly because lower income people pay less rent, are less likely to protest (August 2016), or less general funding compared to previous decades (Tsenkova 2021).  note tsenkova states that less funding to build new housing

```{r 4,fig.cap = "Percentage of apartments in each ward with a score in each group, echo = FALSE }
rent_clean %>%
  mutate(Maintenance_Score = 
           case_when(
             Score <= 50 ~ "Less than 50",
             Score <= 60 ~ "50 to 60",
             Score <= 70 ~ "60 to 70",
             Score <= 80 ~ "70 to 80",
             Score <= 100 ~ "80 or more",
             TRUE ~ "None"
             )
  ) %>%
  ggplot(data=.)+
  theme_minimal()+
  coord_flip()+
  geom_bar(mapping = aes(x = Ward, fill = Maintenance_Score),position="fill")+
  scale_y_continuous(labels = scales::percent)

```

# 2.1 Results

# Discussion

## Societal Impact of these scores

As previously stated, the impact of these scores might lead to an inaccurate rating of social housing apartments or apartments in lower income areas as lower quality, thus perpetuating an even lower view on the people living there, or casting blame on social housing residents for poor conditions instead of the organizations supposed to keep maintenance (cite where poor are blamed for being dirty). Given the importance of these scores in rating apartments and housing choice, there is a need to understand the decisions that go into the scores and how they could be interpreted. It is then possible for these scores to perpetuate harmful views on low income that further progresses their marginalization in society. It also brings into discussion the impact of housing on well being and the needs of the low income that have less than standard or require  lower income housing. There is also the need to discuss the shortcomings of the Toronto Housing Corporation and why they have not been up to date with maintenance of their apartments, as well as public policy that could be formed to improve the maintenance and security of social housing apartments in low income areas. Housing for the lower income is often substandard because they are at lower priority in almost every other facet as life such as law enforcement, schooling, etc. Either because they are often seen as net negative tax payers or that they do not have as much political persuasion (August 2016) as the middle class. This intersects with a large number of poor being racial minorities, who are also often seen as less important in society. Housing is merely one of many ways that systemic discrimination of poor folk and the lack of priority for them.

For social houses that have less security, it is paradoxical as these people often live in more crime ridden areas that needs the extra security, bringing into question if their own lives are considered and the failings of public landlords and the tchc to provide security to private apartment levels. Poor dont have money to protect themselves like rich, so they rely on public government, highlights lack of security for poor in other areas like biased law enforcement.

It has been shown that public housing apartments have lower quality amenities than their private counterparts, and these can have several effects on the residents.
Previous research has shown  that "Aspects of neighbourhoods that have been identified as having an impact on health are: ... “incivilities” such as the presence of graffiti, boarded up houses, garbage accumulation... and broken windows, and the signalling effect of these physical features that no one cares; the perception of neighbourhoods as “safe” and under effective informal social controls...Others have postulated that neighbourhood conditions directly influence habits of privacy, child rearing, house keeping, and study, which can have an indirect effect on health."(Howden-Chapman 2004)

The data showed that Scarborough North, which has a 90% visible minority population with an average income of 78,892 as of 2016, compared to the Toronto average of 100000 (78000 isnt good for an expensive city like Toronto, cite this!) (cite toronto)) has only good rated apartment buildings, an outlier. It is hopeful that average apartment scores for other wards with similar incomes increase in the future. However these low income residents have concerns that a high level of security or good apartments in an otherwise lower income area can make the area more attractive to wealthier people, leading to gentrificiation, which pushes the original low income inhabitants out ((Evans, 2007: 6,7), Hulchanski,2006: Table – Change in Average Individual Income, 1970 to 2000, City of Toronto)(“It's all about power and you have none:” The marginalization of tenant resistance to mixed-income social housing redevelopment in Toronto, Canada,), which is a problem with the already tenuous social housing for these marginalized groups.(Hay,2005: 1)(cite actual)
## Bias and ethical Concerns

There are some areas for possible bias in the data which perpetuate preconceived notions on marginalized groups. As  Rent Safe TO does not release its criteria for scoring, it is unknown what specific attributes they look for while grading. Measurement issues could occur as there is not a way to objectively measure an apartment (i.e it is unclear what distinguishes a rating of 4 from a rating of 5). It is thus possible to then give the an apartment attribute of the same quality a higher score in a wealthier ward/private building with other good amenities, whilst giving apartments in poorer wards/social housing a lower score due to bias against the general area or the social status of residents and not the apartment itself. This perpetuates biases against the lower income and minorities as people gets their previously held beliefs of the poor status of those communities confirmed (cite where people believe minority areas look worse). Lower scores are prevalent in poorer wards and wards with high visible minority populations, and biases against these people and the general "feel" of an area may influence people to rent a more expensive apartment with good scores when there's a cheaper one that was scored wrongly. On the other side, being in a more affluent ward may artificially prop up a mediocre aparmtent's score, thus inflating prices and making it less affordable even though its maintenance is lacking. There is thus a need for a more transparent grading system so that outsiders can cross check and make sure than no historic inequalities perpetuate themselves through scoring biased against poor or social housing folk.

## Collection problem

There are possible data collection problems with this data set. RentSafeTO collects data only for apartment blocks with 3 or more stories and ten or more units(cite), meaning that some buildings that are in dire need of maintenance may be left behind as the organization simply doesn't collect data for them, even though most people may see them as apartments (i.e a three story block of 9 flats). This leaves out a big piece of the full picture on apartment quality, and may also be why some known wealthier wards are getting higher average scores (the low income housing there does not count) and the effect of income inequality as poor people are more likely to live in such types of "apartments" than the wealthy (cite). For apartments that did qualify, the grading is a scheduled event organized with the landlord ahead of time (cite where this is stated on the wbsite). This means that scores presented may not be entirely accurate for general upkeep of the building as landlords who neglect buildings can hurriedly make affairs before the evaluator/enforcement officer arrives. It may thus be necessary for graders to come up semi unannounced to get the true maintenance levels of the building for a more accurate scoring.

## Income
It is important to note that not all living in private apartments are wealthy. In fact, there are many private apartments inhabited by low income people so this disparity in scores between private and social housing may be due to reasons other than income inequality. As stated earlier, less governmental priority to social apartments may be a major reason (Tsenkova 2021).  Average income for axh ward was an important variable as it would put other information into context, as it determines the priority in government spending not just in apartments but in general as lower income areas get less spending but wealthier income are often full of funding in housing, surroundings, and other outside amenities that could have gone into the scoring bias. Unfortunately this variable was not included in the dataset, so information on income had to come from a less up to date source.

(should these be in next steps?)Private housing is now in itself getting harder to afford due to rising housing costs, forcing many so called middle class people into the need for social housing, or forcing them into lower rated apartments. It would be interesting to see the effect of rising prices and the movement of people that would have been considered high income into apartments in lower income wards or social housing, on total maintenance scores and security, to see if the government would then divert funding to these.

## Weaknesses and next steps
We should note that not everyone who lives in private housing is wealthy, and it is important not to overlook those who are struggling to pay rent to access those improved amenities. Not all those who live in private areas or wealthier wards are rich, wealth wards still have poor areas so their own experiences were neglected.

As average income was not provided for the data set, we had to collect external data which was set in 2016 and may not be accurate as it is not current to the modern day.

While the N/As were dispersed randomly throughout the data set, removing them may have caused bias in the data set so our means, medians and analysis of Scores may have been different if the N/A values had been filled.  Also, removing variables even if they were already present in the Score variable may have caused incorrect analysis as there may be some important information in the rejected variables that either brings a new perspective or aids one of the accepted variable in explaining the relationships between apartment type, ward, and maintenance scores.

For next steps we could use the geo-location variable that was omitted for this report to see if the scores for apartments on boundaries between poorer or richer wards had the negative bias of the poor wards or positive bias of the rich.

We could also compare the income and scores of buildings including those that did not have all amenities to further observe if there was inequality as apartments in poor areas tend to have complete absence of some amenities.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References
1.  OECD (2020), “Social housing: A key part of past and future housing policy”, Employment, Labour and Social Affairs Policy Briefs, OECD, Paris, http://oe.cd/social-housing-2020

It's all about power and you have none:” The marginalization of tenant resistance to mixed-income social housing redevelopment in Toronto, Canada

https://www.toronto.ca/city-government/data-research-maps/neighbourhoods-communities/ward-profiles/

Howden-Chapman PHousing standards: a glossary of housing and healthJournal of Epidemiology & Community Health 2004;58:162-168.

https://link.springer.com/chapter/10.1007/978-3-030-69563-7_2

### 2022-02-02 outputs/paper/paper.Rmd
Refined the writing a bit.

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "Housing maintenance beneifts residents as it makes sure all ammenities and appliances function properly, maintains structural integrity, and adds value to homes.
-Previous research has found that private apartments tend to be better maintained than their social housing counterparts, and that housing in richer areas have better ammenities than those in poorer areas.
-We use data from Open data Toronto to further analyze this by breaking down apartment scores and other attributes by Ward and apartment type, using R (R Core Team 2021) and other packages.
+abstract: "Housing maintenance beneifts residents as it ensures proper functioning of ammenities, maintains structural integrity, and adds value to homes.
+However, previous research has found that private apartments tend to be better maintained than their social housing counterparts, and that housing in richer areas have better ammenities than those in poorer areas.
+We use data from Open data Toronto to further analyze this by using plots and tables to break down apartment scores and other attributes by Ward and apartment type, using R (R Core Team 2021) and other packages.
 We find that apartments in higher income areas and privately owned apartments tend to have better maintenance scores than those in lower income wards and public housing respectively.
@@
 
-```{r 4,fig.cap = "Percentage of apartments in each ward with a score in each group, echo = FALSE }
+```{r 4,fig.cap = "Percentage of apartments in each ward with a score in each group", echo = FALSE }
 rent_clean %>%

### 2022-02-02 outputs/paper/paper.Rmd
Updated the R scripts file to read in the dataset, and updated the paper file to add some new writing.

@@
 We then included property type, which states whether an apartment was privately owned, owned by the Toronto Housing Corporation(THCH), or owned by some other non-private institution as our analysis focuses on social/affordable housing.
-The final dataframe included 487 observations with 5 variables.
+The final data frame included 493 observations with 5 variables.
 
@@
   )
+
+rent_clean$Score = strtoi(rent_clean$Score)
+rent_clean$Year_built = as.double(rent_clean$Year_built)
+rent_clean$Security = as.double(rent_clean$Security)
 ```
@@
 )
-```
-```{r}
-rent_clean$Score = strtoi(rent_clean$Score)
-rent_clean$Year_built = as.double(rent_clean$Year_built)
-rent_clean$Security = as.double(rent_clean$Security)
 ```
@@
 rent_clean %>% 
-  summarise(mean=mean(rent_clean$Score),median=median(rent_clean$Score), max = max(rent_clean$Score),min=min(rent_clean$Score))
+  summarise(mean=mean(rent_clean$Score),median=median(rent_clean$Score), max = max(rent_clean$Score),min=min(rent_clean$Score)) %>%
+  knitr::kable(col.names = c("Mean Score", "Median Score", "Maximum Score","Minimum Score"), caption = "Table of Summaries for Apartment Maintenance Scores")
 ```
@@
 High income Wards also tend to have fewer racial minorities, with Toronto saint Paul and Toronto-Danforth having a 72% and 67% white population respectively.
-```{r 1, fig.cap = "Boxplot of mean and median apartment scores by ward", echo = TRUE}
+```{r 1, fig.cap = "Boxplot of median apartment maintenance scores by ward", echo = TRUE}
 rent_clean %>%
@@
   coord_flip()+
-  labs(title = "Wealthier wards tend to have better apartment maintenance than poorer",
+  labs(title = "Wealthier wards tend to have better apartment maintenance than poorer Wards",
        x = "Ward",
@@
 ```
+
+```{r ,fig.cap = "Barplot of Mean Apartment maintenance scores by ward}
+rent_clean %>%
+  ggplot(aes(x = Ward,y = Score)) + 
+  theme_minimal() +
+  geom_bar()
+  coord_flip()+
+  labs(title = "Wealthier wards tend to have better apartment maintenance than poorer",
+       x = "Ward",
+       y = "Average Apartment Score")
+
+```
+
 (Figure @\ref(fig:2)) shows the mean and median apartment scores by type.
@@
   geom_bar(mapping = aes(x = Ward, fill = Maintenance_Score),position="fill")+
+  geom_text(aes(label=Maintenance_Score), vjust=1.6, color="white", size=3.5)+
   scale_y_continuous(labels = scales::percent)
-
+# Add Dark2 here 
 ```

### 2022-02-02 scripts/01-data_cleaning.R

@@ -1,34 +1,44 @@
 #### Preamble ####
-# Purpose: Clean the survey data downloaded from [...UPDATE ME!!!!!]
-# Author: Rohan Alexander [CHANGE THIS TO YOUR NAME!!!!]
-# Data: 3 January 2021
-# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
+# Purpose: Clean the Apartment Maintenance data downloaded from Open Data Toronto 
+# Author: Olaedo Okpareke
+# Data: 2 February 2022
+# Contact: olaedo.okpareke@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-# - Need to have downloaded the ACS data and saved it to inputs/data
-# - Don't forget to gitignore it!
+# - Need to have installed the opendatatoronto and tidyr packages. 
+# - Gitignore it!
 # - Change these to yours
 # Any other information needed?
 
 
 #### Workspace setup ####
 # Use R Projects, not setwd().
-library(haven)
+library(opendatatoronto)
 library(tidyverse)
-# Read in the raw data. 
-raw_data <- readr::read_csv("inputs/data/raw_data.csv"
-                     )
-# Just keep some variables that may be of interest (change 
-# this depending on your interests)
-names(raw_data)
 
-reduced_data <- 
-  raw_data %>% 
-  select(first_col, 
-         second_col)
-rm(raw_data)
-         
+# Read in the raw data.
+apartment_data = list_package_resources("4ef82789-e038-44ef-a478-a8f3590c3eb1") %>%
+  filter(name=="Apartment Building Evaluation") %>%
+  get_resource()
 
+# Cleaning the data by dropping NAs
+apartment_data = na_if(apartment_data, 'N/A')
+
+rent_data <- apartment_data[complete.cases(apartment_data), ]
+
+# Keeping variables of interest
+rent_clean<- rent_data %>% 
+  dplyr::select(Score = SCORE,
+                Ward = WARDNAME,
+                Year_built = YEAR_BUILT,
+                Type= PROPERTY_TYPE,
+                Security = SECURITY
+  )
+
+# Changing variables into the proper variable types
+rent_clean$Score = strtoi(rent_clean$Score)
+rent_clean$Year_built = as.double(rent_clean$Year_built)
+rent_clean$Security = as.double(rent_clean$Security)
 #### What's next? ####
 
 

### 2022-02-02 outputs/paper/paper.Rmd
Added some references to the bib file, added new writing to the paper, deleted some words in the data r script.

@@
   - Another author
-thanks: "Code and data are available at: github repo. What addres do i put?."
+thanks: "Code and data are available at: https://github.com/OlaedoOkpareke/starter_folder."
 date: "`r format(Sys.time(), '%d %B %Y')`"
 abstract: "Housing maintenance beneifts residents as it ensures proper functioning of ammenities, maintains structural integrity, and adds value to homes.
-However, previous research has found that private apartments tend to be better maintained than their social housing counterparts, and that housing in richer areas have better ammenities than those in poorer areas.
+However, previous research has found that priority in maintenance for public housing apartments and lower income apartments is little compared to their more privileged counterparts.
 We use data from Open data Toronto to further analyze this by using plots and tables to break down apartment scores and other attributes by Ward and apartment type, using R (R Core Team 2021) and other packages.
-We find that apartments in higher income areas and privately owned apartments tend to have better maintenance scores than those in lower income wards and public housing respectively.
+We find that apartments in higher income areas and privately owned apartments tend to have better evaluation scores than those in lower income wards and public housing respectively.
 This data shows the inequality in standards for housing in Toronto, as well as the need for better government accountability and housing justice.
@@
 Toronto is a large Canadian city divided into 25 wards.
-In order to make sure building maintenance standards are upheld and to allow apartment residents to look up safety scores of their apartments, Rent Safe TO has provides a dataset which, according to a set criteria not revealed to public, rates apartment attributes according to a 5 point system, and then uses a formula to assign them an overall score over 100.
+In order to make sure building maintenance standards are upheld and to allow apartment residents to look up safety scores of their apartments, Rent Safe TO has provides an apartment evaluation dataset dataset.
+According to a set criteria not revealed to public, this dataset rates apartment attributes according to a 5 point system, and then uses a formula to assign them an overall score over 100.
 In this report, we are interested in how apartment maintenance scores have changed over the years, and how attributes such as relative wealth of the ward and the type of apartment affects the total score given to an apartment.
@@
 Figures and tables were created with ggplot2 (Wickham 2016) and kableextra (Zhu 2021).
-We created tables that first showed the data and important summary statistics, then went into showing summaries for area and apartment types, as well as box plots, linear model graphs and percentage bar charts that highlighted inequality in housing scores.
+We created tables that first showed the data and important summary statistics, then went into showing the data for area and apartment types, as well as box plots, linear model graphs and percentage bar charts, which highlighted inequality in housing evaluation scores.
 
@@
 Apartments in some wealthier wards also have a higher average maintenance score than those in poorer wards.
+Private apartments also tends to have better security than their social housing counterparts.
 According to (OECD 2020) and (Desroches and Poland, 2021), these disparities are due to a decline in public investment.
@@
 1994, pg 146).
-This data and the statistics shows that vulnerable populations such as low income and minorities suffer systemic marginalization in facets of society, even in the housing sector as well maintained, secure housing is still unattainable for the city' most vulnerable, questioning the extent of the achievements of social housing.
-There are also concerns on gentrification, bias in grading, and ...
+This data and the statistics shows that vulnerable populations such as low income and minorities suffer systemic marginalization in facets of society, even in the housing sector as well maintained, secure housing is still unattainable for the city's most vulnerable, questioning the extent of the achievements of social housing.
+There are also concerns on the bias of the grading, and the impact of these scores on gentrification.
 
@@
 We also excluded the ward number variable as the ward names variable were easier to identify.
-We included both year built as we wanted to see the change in housing maintenance between older and newer apartments.
+We included year built as we wanted to see the change in apartment evaluation between older and newer apartments.
 We included score as it was the main determinant of maintenance standards, as well as security to observe the the differences in security between different groups of apartments, and we wanted to see its direct effect.
@@
 A preview of the data, (Table \@ref(tab:1)) using kableexrra(Zhu 2021) is shown below.
-```{r 1}
+```{r 1, fig.cap = "Table of the first few rows of the data"}
 x = head(rent_clean) %>%
@@
 We can see from the table that the average building has good standards and security, but is a bit old.
-The average apartment has a better score than the median apartment, indicating that there are some apartments that are far above the average.However the mean security is lower the median, indicating some very low security apartments.
-Policy is needed to improve security for these apartments, and the disrepancies that some scores are far higher than most are put into question.
+The average apartment has a better score than the median apartment, indicating that there are some apartments that are far above the average.
+However the mean security is lower the median, indicating some very low security apartments.
+Policy is needed to improve security for these apartments, and the observation that some scores are far higher than most are more thoroughly scrutinized later in the report.
 
-```{r 2}
+```{r 2, fig.cap = "Table with summaries of the Score and Security for all apartments"}
 rent_clean %>% 
-  summarise(mean=mean(rent_clean$Score),median=median(rent_clean$Score), max = max(rent_clean$Score),min=min(rent_clean$Score)) %>%
-  knitr::kable(col.names = c("Mean Score", "Median Score", "Maximum Score","Minimum Score"), caption = "Table of Summaries for Apartment Maintenance Scores")
+  summarise(mean=mean(rent_clean$Score),median=median(rent_clean$Score), mean1 = mean(rent_clean$Security),median2=median(rent_clean$Security)) %>%
+  knitr::kable(col.names = c("Mean Score", "Median Score", "Mean Security","Median Security"), caption = "Table of Summaries for Apartment Maintenance Scores")
 ```
 
-(Figure @\ref(fig: 1)), using ggplot2(Wickham 2016), shows the average maintenance scores by ward.
-Cross referencing the income and demographics of each ward using (City of Toronto 2022) , we see that the higher income wards such as Toronto Saint Paul's and Toronto-Danforth (155,470 and 101,323 average yearly income respectively 72% white and 67% white respectively as of 2016) tend to have both higher means and medians than lower income wards, such as Davenport and Humber River Black Creek (80,807 and 65,458 average yearly income as of 2016).
+(Figure @\ref(fig: 1)), using ggplot2 (Wickham 2016), shows the average maintenance scores by ward.
+Cross referencing the income and demographics of each ward using (City of Toronto 2022) , we see that the higher income wards such as Toronto Saint Paul's and Toronto-Danforth (155,470 and 101,323 average yearly income respectively 72% white and 67% white respectively as of 2016) tend to have both higher means and medians than lower income wards, such as Davenport and Humber River-Black Creek (80,807 and 65,458 average yearly income as of 2016).
 An exception seems to be Scarborough North which has a low yearly average income and high racial minority population but a high average apartment maintenance score.
-However note that the median is very small, implies that the sample in this population may not be representative of the true state of apartments in this ward.
-Although there are several outliers Scores for each ward, in general apartment housing is better in richer wards than in poorer wards.
-High income Wards also tend to have fewer racial minorities, with Toronto saint Paul and Toronto-Danforth having a 72% and 67% white population respectively.
+However note that the range is very small, implying that there are very few samples, so they may not be representative of the true state of apartments in this ward.
+Although there are several outlier Scores for each ward, in general apartment housing is better in richer wards than in poorer wards.
+High income Wards also tend to have fewer racial minorities, with Toronto saint Paul and Toronto-Danforth having a 72% and 67% non-racial minority population respectively.
 ```{r 1, fig.cap = "Boxplot of median apartment maintenance scores by ward", echo = TRUE}
@@
   coord_flip()+
-  labs(title = "Wealthier wards tend to have better apartment maintenance than poorer Wards",
+  labs(title = "Wealthier wards have better scores than poorer Wards",
        x = "Ward",
@@
 The spread of quality seems to be equal for all three housing types, so even accounting for variation private housing is consistently better than their public counterparts.
-This shows a disparity in priority for apartment maintenance and upkeep as well as general housing inequality, and shows that government are not on top of social housing affairs, as those in social housing owned by the TCHC also deserve high average maintenance taking care of their social housing.
-However it should be noted that (non THCH owned) social housing maintenance is not very far behind private housing, showing a just upkeep of housing.
-However according to (), people in social housing often have to take maintenance and upkeep into their own hands.
-```{r 2,fig.cap = "Boxplot of mean and median apartment scores by Apartment Type", echo = TRUE }
+This shows a disparity in priority for apartment maintenance and upkeep, as well as general housing inequality as those in social housing owned by the TCHC also deserve high average maintenance but do not, and reveals that the government are not on top of taking care of social housing.
+However it should be noted that (non TCHC owned) social housing maintenance is not very far behind private housing, showing a just upkeep of housing.
+Nevertheless according to (), people in social housing often have to take maintenance and upkeep into their own hands.
+```{r 2,fig.cap = "Boxplot of median apartment scores by Apartment Type", echo = TRUE }
 rent_clean %>%
@@
   geom_boxplot() +
-  labs(title = "Private Housing has better scores than social housing of thch and not thch",
+  labs(title = "Private Housing has better scores than social and TCHC housing ",
        x = "Apartment Type",
@@
 (Table \@ref(tab:3))  shows the differences in security between each housing type.
-We see that overall the average maintenance scores are good for all types of apartment.
-However the median shows that both THCH are lower than that of the private buildings, and their average scores were driven up by  a few very good apartments.
+We see that overall the average secuirty scores are good for all types of apartment.
+However the median shows that both TCHC owned and non TCHC owned social buildings have less stellar security than that of the private buildings, so their average scores were driven up by  a few very good apartments.
 This means that we can expect that a large number of social housing (which houses mainly low income populations) do not have the best security, which is counter-intuitive as poorer areas where this housing is often located are often in higher need of security due to correlation with crime rates.
@@
 
-(Figure @\ref(fig:3)) shows the general trend of apartment maintenance scores for the 3 types according to the year they were built.
-Even though for all types newer buildings have better maintenance Scores than their older counterparts for all three types of apartment, private housing still gets consistently higher scores than the other two types for almost all the years.
+(Figure @\ref(fig:3)) shows the general trend of apartment maintenance scores for all types according to the year they were built.
+Even though for all three types newer buildings have better maintenance Scores than their older counterparts, private housing still gets consistently higher scores than the other two types for almost all the years.
 Most private apartments were built between the 1960s to 1970s while social housing was built between the 1980s to 2000s, so even the newer age of the social housing buildings is not enough to protect against lower rates of maintenance.
-This shows either the persistence in lack of government and public maintenance or persistent bias in scoring, which all tie into income inequality.
+This shows either the persistence in lack of government and public maintenance or persistent bias in scoring, which both tie into income inequality.
 It should be noted that the trend for private apartments may be heavily influenced by the points at the far right end, and that many private apartments also have low scores.
@@
 
-(Figure @\ref(fig:4)) shows that the wealthier wards have a larger percentage of their apartmentd scoring above 70 compared to lower income wards.
+(Figure @\ref(fig:4)) shows that the wealthier wards have a larger percentage of their apartments scoring above 70 compared to lower income wards.
 Some of the poorer wards such as Humber River Black creek or Davenport do not have very highly rated apartments at all.
 The racial minority population in these two wards as of 2016 was 74% and 43% respectively (City of Toronto 2022).In 2016 the income for these wards was 65,548 and 80,807 dollars respectively while in the richer wards like Toronto Saint Paul and Etobicoke Centre it was 155,470 and 128,448 (City of Toronto 2022).
-Many other wards follow the same trend, with the wards with higher income having larger percentages of their apartments being above 70, with some lower income wards even having some percentage of their apartments have a score less than 50.
-This is another way to see the inequality in housing standards, and communicates that housing maintenance disproportionally caters to higher income areas, possibly because lower income people pay less rent, are less likely to protest (August 2016), or less general funding compared to previous decades (Tsenkova 2021).
+Many other wards follow the same trend, with the wards with higher income having larger percentages of their apartments scoring above 70, with some lower income wards even having some percentage of their apartments have a score less than 50.
+This is another way to see the inequality in housing standards, and communicates that housing maintenance disproportionally caters to higher income areas.
+This is possibly because lower income people pay less rent, are less likely to protest (August 2016), or less general funding compared to previous decades (Tsenkova 2021).
 note tsenkova states that less funding to build new housing
 
-```{r 4,fig.cap = "Percentage of apartments in each ward with a score in each group", echo = FALSE }
+```{r 4,fig.cap = "Percentage of apartments in each ward with that have a score in a certain group", echo = FALSE }
 rent_clean %>%
@@
 
-As previously stated, the impact of these scores might lead to an inaccurate rating of social housing apartments or apartments in lower income areas as lower quality, thus perpetuating an even lower view on the people living there, or casting blame on social housing residents for poor conditions instead of the organizations supposed to keep maintenance (cite where poor are blamed for being dirty).
+As previously stated, the impact of these scores might lead to an inaccurate rating of social housing apartments or apartments in lower income areas as lower quality, thus perpetuating harmful views on low income people that further progress their marginalization in society by casting blame on social housing residents for poor conditions instead of the organizations supposed to keep maintenance (cite where poor are blamed for being dirty).
 Given the importance of these scores in rating apartments and housing choice, there is a need to understand the decisions that go into the scores and how they could be interpreted.
-It is then possible for these scores to perpetuate harmful views on low income that further progresses their marginalization in society.
-It also brings into discussion the impact of housing on well being and the needs of the low income that have less than standard or require  lower income housing.
-There is also the need to discuss the shortcomings of the Toronto Housing Corporation and why they have not been up to date with maintenance of their apartments, as well as public policy that could be formed to improve the maintenance and security of social housing apartments in low income areas.
-Housing for the lower income is often substandard because they are at lower priority in almost every other facet as life such as law enforcement, schooling, etc.
-Either because they are often seen as net negative tax payers or that they do not have as much political persuasion (August 2016) as the middle class.
-This intersects with a large number of poor being racial minorities, who are also often seen as less important in society.
-Housing is merely one of many ways that systemic discrimination of poor folk and the lack of priority for them.
 
-For social houses that have less security, it is paradoxical as these people often live in more crime ridden areas that needs the extra security, bringing into question if their own lives are considered and the failings of public landlords and the tchc to provide security to private apartment levels.
-Poor dont have money to protect themselves like rich, so they rely on public government, highlights lack of security for poor in other areas like biased law enforcement.
-
+It also brings into discussion the impact of housing on well being and the needs of the low income that have less than standard housing or require  lower income housing.
 It has been shown that public housing apartments have lower quality amenities than their private counterparts, and these can have several effects on the residents.
@@
 
-The data showed that Scarborough North, which has a 90% visible minority population with an average income of 78,892 as of 2016, compared to the Toronto average of 100000 (78000 isnt good for an expensive city like Toronto, cite this!) (cite toronto)) has only good rated apartment buildings, an outlier.
+There is also the need to discuss the shortcomings of the Toronto Housing Corporation and why they have not been up to date with maintenance of their apartments, as well as public policy that could be formed to improve the maintenance and security of social housing apartments in low income areas.
+Housing for the lower income is often substandard because they are at lower priority in almost every other facet as life such as law enforcement, schooling, etc.
+Either because they are often seen as net negative tax payers or that they do not have as much political persuasion (August 2016) as the middle class.
+This intersects with a large number of poor being racial minorities, who are also often seen as less important in society.
+"Housing is merely one of many ways that systemic discrimination of poor folk and the lack of priority for them is shown.
+Lack of affordable or poor quality housing exacerbates challenges associated with low income, education, literacy, physical and mental disabilities, race, ethnicity and colour" (Carter, 2004: 25).
+
+Regarding social housing apartments that have less security, it is paradoxical as these people often live in more crime ridden areas that needs the extra security, bringing into question if their own lives are considered and the failings of public landlords and the TCHC to provide security akin to those of private apartment.
+Lower income people often do not have money to protect themselves like rich, so they rely on publicly provided assistance.
+(cite?) This also highlights lack of security for poor in other areas like biased law enforcement.
+(cite)
+
+The data showed that Scarborough North, which has a 90% visible minority population with an average income of 78,892 as of 2016, compared to the Toronto average of 100000 (cite toronto)) only has highly scored apartment buildings, an outlier.
 It is hopeful that average apartment scores for other wards with similar incomes increase in the future.
-However these low income residents have concerns that a high level of security or good apartments in an otherwise lower income area can make the area more attractive to wealthier people, leading to gentrificiation, which pushes the original low income inhabitants out ((Evans, 2007: 6,7), Hulchanski,2006: Table – Change in Average Individual Income, 1970 to 2000, City of Toronto)(“It's all about power and you have none:” The marginalization of tenant resistance to mixed-income social housing redevelopment in Toronto, Canada,), which is a problem with the already tenuous social housing for these marginalized groups.(Hay,2005: 1)(cite actual)
+However low income residents have concerns that a high level of security or good apartments in an otherwise lower income area can make the area more attractive to wealthier people, leading to gentrification, which pushes the original low income inhabitants out ((Evans, 2007: 6,7), Hulchanski,2006: Table – Change in Average Individual Income, 1970 to 2000, City of Toronto)(“It's all about power and you have none:” The marginalization of tenant resistance to mixed-income social housing redevelopment in Toronto, Canada,), which is a problem with the already tenuous social housing for these marginalized groups.(Hay,2005: 1)(cite actual)
+
 ## Bias and ethical Concerns
@@
 This leaves out a big piece of the full picture on apartment quality, and may also be why some known wealthier wards are getting higher average scores (the low income housing there does not count) and the effect of income inequality as poor people are more likely to live in such types of "apartments" than the wealthy (cite).
-For apartments that did qualify, the grading is a scheduled event organized with the landlord ahead of time (cite where this is stated on the wbsite).
+For apartments that did qualify, the grading is a scheduled event organized with the landlord ahead of time (cite where this is stated on the website).
 This means that scores presented may not be entirely accurate for general upkeep of the building as landlords who neglect buildings can hurriedly make affairs before the evaluator/enforcement officer arrives.
@@
 As stated earlier, less governmental priority to social apartments may be a major reason (Tsenkova 2021).
-Average income for axh ward was an important variable as it would put other information into context, as it determines the priority in government spending not just in apartments but in general as lower income areas get less spending but wealthier income are often full of funding in housing, surroundings, and other outside amenities that could have gone into the scoring bias.
+The average income for each ward was an important variable as it would put other information into context, as it determines the priority in government spending not just in apartments but in general as lower income areas get less spending but wealthier income are often full of funding in housing, surroundings, and other outside amenities that could have gone into the scoring bias.
 Unfortunately this variable was not included in the dataset, so information on income had to come from a less up to date source.

### 2022-02-02 scripts/01-data_cleaning.R

@@ -39,7 +39,7 @@ rent_clean<- rent_data %>%
 rent_clean$Score = strtoi(rent_clean$Score)
 rent_clean$Year_built = as.double(rent_clean$Year_built)
 rent_clean$Security = as.double(rent_clean$Security)
-#### What's next? ####
+
 
 
 

### 2022-02-04 outputs/paper/paper.Rmd
Added text to the readme file.
Made several changes and modifications to these files.

@@
 title: "Income Inequality is revealed in Torontonian Apartment Maintenance"
-subtitle: "My subtitle if needed"
 author: 
   - Olaedo Okpareke
-  - Another author
 thanks: "Code and data are available at: https://github.com/OlaedoOkpareke/starter_folder."
@@
 knitr::opts_chunk$set(echo = TRUE)
-
 library(tidyverse)
-install.packages("opendatatoronto")
+# install.packages("opendatatoronto")
 library(opendatatoronto)
@@
 ```{r}
-apartment_data = list_package_resources("4ef82789-e038-44ef-a478-a8f3590c3eb1") %>%
-  filter(name=="Apartment Building Evaluation") %>%
-  get_resource()
-
-apartment_data = na_if(apartment_data, 'N/A')
-
-rent_data <- apartment_data[complete.cases(apartment_data), ]
-
-# Try  to filter by year evaluated and see if it gets what I want
-rent_clean<- rent_data %>% 
-  dplyr::select(Score = SCORE,
-                Ward = WARDNAME,
-                Year_built = YEAR_BUILT,
-                Type= PROPERTY_TYPE,
-                Security = SECURITY
-  )
-
-rent_clean$Score = strtoi(rent_clean$Score)
-rent_clean$Year_built = as.double(rent_clean$Year_built)
-rent_clean$Security = as.double(rent_clean$Security)
-```
-
-```{r}
-rent_clean %>%
-  type_convert(
-  rent_clean,
-  col_types = cols(
-    Score = col_integer(),
-    Ward = col_character(),
-    Year_built = col_double(),
-    Type = col_character(),
-    Security = col_double()
-  ),
-  na = c("", "N/A"),
-  guess_integer = FALSE
-)
+rent_clean = read_csv(here::here("inputs/data/rent_clean.csv"))
 ```
@@
 It shows a measure of income inequality in housing, and has possible implications for public policy.
-A preview of the data, (Table \@ref(tab:1)) using kableexrra(Zhu 2021) is shown below.
-```{r 1, fig.cap = "Table of the first few rows of the data"}
+A preview of the data, (Table \@ref(tab:tableone)) and figure 1 using kableexrra(Zhu 2021) and ggplot2 (Wickham...) is shown below.
+
+```{r tableone, fig.cap = "Table of the first few rows of the data"}
 x = head(rent_clean) %>%
@@
 ```
+```{r figone, fig.cap = "Table with summaries of the Score and Security for all apartments"}
+ggplot(rent_clean, aes(x=Score)) +
+  geom_histogram(fill="black")+
+  theme_minimal()
+```
 
-(Table \@ref(tab:2)) shows the the summary statistics for the buildings.
+(Figure \@ref(fig:figone)) shows the the summary statistics for the buildings.
+We see that the data seems to be unimodal and normally distributed.
+There is an outlier which skews the data.
 We can see from the table that the average building has good standards and security, but is a bit old.
@@
 
-```{r 2, fig.cap = "Table with summaries of the Score and Security for all apartments"}
-rent_clean %>% 
-  summarise(mean=mean(rent_clean$Score),median=median(rent_clean$Score), mean1 = mean(rent_clean$Security),median2=median(rent_clean$Security)) %>%
-  knitr::kable(col.names = c("Mean Score", "Median Score", "Mean Security","Median Security"), caption = "Table of Summaries for Apartment Maintenance Scores")
-```
-
-(Figure @\ref(fig: 1)), using ggplot2 (Wickham 2016), shows the average maintenance scores by ward.
+(Figure \@ref(fig:figtwo)), using ggplot2 (Wickham 2016), shows the average maintenance scores by ward.
 Cross referencing the income and demographics of each ward using (City of Toronto 2022) , we see that the higher income wards such as Toronto Saint Paul's and Toronto-Danforth (155,470 and 101,323 average yearly income respectively 72% white and 67% white respectively as of 2016) tend to have both higher means and medians than lower income wards, such as Davenport and Humber River-Black Creek (80,807 and 65,458 average yearly income as of 2016).
@@
 High income Wards also tend to have fewer racial minorities, with Toronto saint Paul and Toronto-Danforth having a 72% and 67% non-racial minority population respectively.
-```{r 1, fig.cap = "Boxplot of median apartment maintenance scores by ward", echo = TRUE}
+
+```{r figtwo, fig.cap = "Boxplot of median apartment maintenance scores by ward", echo = TRUE}
+
 rent_clean %>%
@@
 
-```{r ,fig.cap = "Barplot of Mean Apartment maintenance scores by ward}
-rent_clean %>%
-  ggplot(aes(x = Ward,y = Score)) + 
-  theme_minimal() +
-  geom_bar()
-  coord_flip()+
-  labs(title = "Wealthier wards tend to have better apartment maintenance than poorer",
-       x = "Ward",
-       y = "Average Apartment Score")
-
-```
-
-(Figure @\ref(fig:2)) shows the mean and median apartment scores by type.
+(Figure \@ref(fig:figthree)) shows the mean and median apartment scores by type.
 The results show that private buildings have the best average maintenance quality, and those from the Toronto housing Corporation have the lowest average.
@@
 Nevertheless according to (), people in social housing often have to take maintenance and upkeep into their own hands.
-```{r 2,fig.cap = "Boxplot of median apartment scores by Apartment Type", echo = TRUE }
+
+```{r figthree, fig.cap = "Boxplot of median apartment scores by Apartment Type", echo = TRUE }
 rent_clean %>%
@@
 ```
-(Table \@ref(tab:3))  shows the differences in security between each housing type.
+
+(Table \@ref(tab:tabtwo))  shows the differences in security between each housing type.
 We see that overall the average secuirty scores are good for all types of apartment.
@@
 This could be alleviated with multi-income neighbourhoods, which places many different types of housing that can accomodate different income groups, but NIMBY-ism has slowed the progress of these neighbourhoods.
-```{r}
+```{r tabtwo}
 rent_clean %>%
-  group_by(Type) %>%# We want to know the occupancy by month
+  group_by(Type) %>%# 
   summarize(Mean = mean(Security),Median = median(Security)) %>%
-  knitr::kable()
+  knitr::kable(caption = "Table of Mean and Median security for each group") 
 ```
 
-(Figure @\ref(fig:3)) shows the general trend of apartment maintenance scores for all types according to the year they were built.
+(Figure @\ref(fig:figfour)) shows the general trend of apartment maintenance scores for all types according to the year they were built.
 Even though for all three types newer buildings have better maintenance Scores than their older counterparts, private housing still gets consistently higher scores than the other two types for almost all the years.
@@
 We also removed an outlier, 1910 as it affected the accuracy of the trend for private apartments.
-```{r 3,fig.cap = "Scatterplots and regression line of trend of apartment scores by type over 1950 - 2021", echo = FALSE }
+```{r figfour,fig.cap = "Scatterplots and regression line of trend of apartment scores by type over 1950 - 2021", echo = FALSE }
 rent_clean %>%
@@
        y = "Score",
-       color = "Type") +
-  scale_color_brewer(palette = "Dark2")
+       color = "Type") 
 ```
 
-(Figure @\ref(fig:4)) shows that the wealthier wards have a larger percentage of their apartments scoring above 70 compared to lower income wards.
+(Figure @\ref(fig:figfive)) shows that the wealthier wards have a larger percentage of their apartments scoring above 70 compared to lower income wards.
 Some of the poorer wards such as Humber River Black creek or Davenport do not have very highly rated apartments at all.
@@
 
-```{r 4,fig.cap = "Percentage of apartments in each ward with that have a score in a certain group", echo = FALSE }
+```{r figfive,fig.cap = "Percentage of apartments in each ward with that have a score in a certain group", echo = FALSE }
 rent_clean %>%
@@
   geom_bar(mapping = aes(x = Ward, fill = Maintenance_Score),position="fill")+
-  geom_text(aes(label=Maintenance_Score), vjust=1.6, color="white", size=3.5)+
   scale_y_continuous(labels = scales::percent)
@@
 The average income for each ward was an important variable as it would put other information into context, as it determines the priority in government spending not just in apartments but in general as lower income areas get less spending but wealthier income are often full of funding in housing, surroundings, and other outside amenities that could have gone into the scoring bias.
-Unfortunately this variable was not included in the dataset, so information on income had to come from a less up to date source.
+Unfortunately this variable was not included in the data set, so information on income had to come from a less up to date source.
 
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional details
-
-\newpage
-
 # References

### 2022-02-04 scripts/01-data_cleaning.R

@@ -40,7 +40,9 @@ rent_clean$Score = strtoi(rent_clean$Score)
 rent_clean$Year_built = as.double(rent_clean$Year_built)
 rent_clean$Security = as.double(rent_clean$Security)
 
+# create data set from csv
 
+write_csv(rent_clean, "inputs/data/rent_clean.csv")
 
 
          

### 2022-02-05 outputs/paper/paper.Rmd
Correcting text, adding citations.

@@
 title: "Income Inequality is revealed in Torontonian Apartment Maintenance"
-author: 
-  - Olaedo Okpareke
-thanks: "Code and data are available at: https://github.com/OlaedoOkpareke/starter_folder."
+author: "Olaedo Okpareke"
+thanks: 'Code and data are available at: https://github.com/OlaedoOkpareke/starter_folder'
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "Housing maintenance beneifts residents as it ensures proper functioning of ammenities, maintains structural integrity, and adds value to homes.
-However, previous research has found that priority in maintenance for public housing apartments and lower income apartments is little compared to their more privileged counterparts.
-We use data from Open data Toronto to further analyze this by using plots and tables to break down apartment scores and other attributes by Ward and apartment type, using R (R Core Team 2021) and other packages.
+abstract: "Housing maintenance benefits residents as it ensures proper functioning of ammenities, maintains structural integrity, and adds value to homes.
+However, previous research has found that priority in maintenance for public housing apartments and lower income apartments is little compared to their more privileged counterparts.We use data from Open data Toronto to further analyze this by using plots and tables to break down apartment scores and other attributes by Ward and apartment type,using R (R Core Team 2021) and other packages.
 We find that apartments in higher income areas and privately owned apartments tend to have better evaluation scores than those in lower income wards and public housing respectively.
-This data shows the inequality in standards for housing in Toronto, as well as the need for better government accountability and housing justice.
-Improving the gap in aparment quality helps improve the life of poor and marginalized people."
+This report shows the inequality in standards for housing in Toronto, as well as the need for better government accountability and housing justice.
+Improving the gap in aparment quality helps improve the lives of poor and marginalized people."
 output:
@@
 ---
-BOXPLOTS TALK ABOYT MEDIANS< NOT MEANS. PUT THE TABLE FOR MEAN
+
 ```{r setup, include=FALSE}
@@
 
-# 1. Introduction
+# Introduction
 Toronto is a large Canadian city divided into 25 wards.
-In order to make sure building maintenance standards are upheld and to allow apartment residents to look up safety scores of their apartments, Rent Safe TO has provides an apartment evaluation dataset dataset.
-According to a set criteria not revealed to public, this dataset rates apartment attributes according to a 5 point system, and then uses a formula to assign them an overall score over 100.
-In this report, we are interested in how apartment maintenance scores have changed over the years, and how attributes such as relative wealth of the ward and the type of apartment affects the total score given to an apartment.
+In order to make sure building maintenance standards are upheld and to allow apartment residents to look up safety scores of their apartments, Rent Safe TO has provides an apartment evaluation data set data set.
+According to a set criteria not revealed to public, this data set rates apartment attributes according to a 5 point system, and then uses a formula to assign them an overall score over 100.
+In this report, we are interested in how attributes such as relative wealth of the ward and the type of apartment affects the total maintenance score of an apartment, and how these scores have changed over the years.
 
-We obtained the dataset from the City of Toronto Open Data Portal [the address of this book](https://open.toronto.ca/dataset/apartment-building-evaluation/) with `opendatatoronto` (Gelfand 2020), and cleaned and analyzed the dataset using R (R Core Team 2021),tidyverse (Wickham et al.
-2019) and dplyr (Wickham et al.
-2021) packages.
+We obtained the data set from the City of Toronto Open Data Portal (https://open.toronto.ca/dataset/apartment-building-evaluation/) with `opendatatoronto` (Gelfand 2020), and cleaned and analyzed the data set using R (R Core Team 2021),tidyverse (Wickham et al.
+2019) and dplyr (@citetidyverse) packages.
 Figures and tables were created with ggplot2 (Wickham 2016) and kableextra (Zhu 2021).
-We created tables that first showed the data and important summary statistics, then went into showing the data for area and apartment types, as well as box plots, linear model graphs and percentage bar charts, which highlighted inequality in housing evaluation scores.
+We created tables that first showed the data and important summary statistics, then went into showing the data for area and apartment types using box plots, linear model graphs and percentage bar charts, which highlighted inequality in housing evaluation scores.
 
@@
 Apartments in some wealthier wards also have a higher average maintenance score than those in poorer wards.
-Private apartments also tends to have better security than their social housing counterparts.
+Private apartments also tend to have better security than their social housing counterparts.
 According to (OECD 2020) and (Desroches and Poland, 2021), these disparities are due to a decline in public investment.
@@
 
-# 2 Data
+# Data
 The data for this report was gotten from the Apartment Building Evaluation (Rent Safe TO 2022) from the Toronto Open Data Portal (city of Toronto 2022) using the R package opendatatoronto (Gelfand 2020).
 This data is compiled by RentSafeTO and was last refreshed on January 22nd, 2022.
-Rent Safe TO collects data on apartments in the GTA that are 3 or more stories or had 10 more units, and response is mandatory for qualifying apartments, i.e the population.
-The raw data includes 9758 apartment blocks from the Greater Toronto Area and their categorical characteristics (40 variables in total) such as location and Type of property, as well as numeric characteristics such as Security,ammenities and state of common areas.
+Rent Safe TO collects data on apartments in the GTA that are 3 or more stories or had 10 more units (cite this), and response is mandatory for qualifying apartments, i.e the population.
+The raw data includes 9758 apartment blocks from the Greater Toronto Area and their categorical characteristics (40 variables in total) such as location and Type of property, as well as numeric characteristics such as Security,amenities and state of common areas.
 It also includes an overall score over 100 for each apartment block.
 
-For data cleaning, we removed all rows with missing values as they indicated that the apartment did not have a certain room, ammenity, or characteristic, but we wanted an "all things being equal" analysis, the assumption that attributes wise, they were alike.
+For data cleaning, our intended sample was apartments that included all available rooms, amenities, and other characteristics.
+This was because we wanted an "all things being equal" analysis, the assumption that amenities wise, the apartments were alike.
 That is, we wanted to compare apartments that a had poor/ lacking characteristic, not no characteristic at all.
-It was also helpful to see the full picture so we could see if there was bias in grading instead of it not being graded at all.
-The missing values seemed to be randomly dispersed about the dataset so removing them would not cause much deviation.
+We removed all rows with missing values as they indicated one or more of the previous criteria was missing.
+It was also helpful to see the full picture to observe the possibility of bias in grading instead of no grading at all.
+The missing values seemed to be randomly dispersed about the data set so removing them would not cause much deviation.
 
@@
 we first removed _id and RSN as they were merely ID variables.
-We also excluded the ward number variable as the ward names variable were easier to identify.
-We included year built as we wanted to see the change in apartment evaluation between older and newer apartments.
-We included score as it was the main determinant of maintenance standards, as well as security to observe the the differences in security between different groups of apartments, and we wanted to see its direct effect.
-We then included property type, which states whether an apartment was privately owned, owned by the Toronto Housing Corporation(THCH), or owned by some other non-private institution as our analysis focuses on social/affordable housing.
+We also excluded the ward number variable as the ward names variable (Ward) was easier to use for identification.
+We included year built as we wanted to see the changes in apartment evaluation between older and newer apartments.
+We included score as it was the main determinant of maintenance standards.
+We also included Security to observe the the direct relationship and differences in security between different groups of apartments.
+We then included property type, which states whether an apartment was privately owned, owned by the Toronto Housing Corporation (TCHC), or owned by some other non-private institution as our analysis focuses on social/affordable housing.
 The final data frame included 493 observations with 5 variables.
 
-```{r}
+```{r,include=FALSE}
 rent_clean = read_csv(here::here("inputs/data/rent_clean.csv"))
@@
 
-We are interested in the relationships between property Type and Ward on Score and security, as well as the trend of Score over time.
-It shows a measure of income inequality in housing, and has possible implications for public policy.
-A preview of the data, (Table \@ref(tab:tableone)) and figure 1 using kableexrra(Zhu 2021) and ggplot2 (Wickham...) is shown below.
+We are interested in the relationships between apartment Type and Ward on Score and Security, as well as the trend of Score over time.
+These relationships show a measure of income inequality in housing, and have possible implications for public policy.
+A preview of the data, (Table \@ref(tab:tabone)), (Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) using kableextra (Zhu 2021) and ggplot2 (Wickham 2016) respectively is shown below.
 
-```{r tableone, fig.cap = "Table of the first few rows of the data"}
+```{r tabone, fig.cap = "Table of the first few rows of the data", echo = FALSE}
+
 x = head(rent_clean) %>%
@@
 ```
-```{r figone, fig.cap = "Table with summaries of the Score and Security for all apartments"}
+
+(Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) show the spread and distribution of apartments scores and their security levels.
+We see that the data for Scores and Age seems to be unimodal and normally distributed.
+There is an outlier of a very low apartment score, but it is not enough to skew the data to the left.
+We can see from the figures that the average building has good maintenance and security, but is a bit old.
+The median apartment score seems to be the same or a slightly above the mean apartment score, as there seems to be more apartments on the lower ends than there are on the higher end.
+This indicates there are many below average apartments.
+As there are quite of lower security apartments, it seems that the mean Security will be lower than the median.
+some very low security apartments.
+Policy is needed to improve security for these apartments, and the observation that some evaluation Scores are quite lower than most are more thoroughly scrutinized later in the report.
+
+```{r figone,out.width="60%", fig.cap = "Histogram of the Distribution of Apartment Scores", echo= FALSE}
+par(mfrow=c(2,2))
 ggplot(rent_clean, aes(x=Score)) +
-  geom_histogram(fill="black")+
-  theme_minimal()
+  geom_histogram(color = "black", fill="white")+
+  theme_minimal()+
+  xlab("Evaluation Score")
 ```
 
-(Figure \@ref(fig:figone)) shows the the summary statistics for the buildings.
-We see that the data seems to be unimodal and normally distributed.
-There is an outlier which skews the data.
-We can see from the table that the average building has good standards and security, but is a bit old.
-The average apartment has a better score than the median apartment, indicating that there are some apartments that are far above the average.
-However the mean security is lower the median, indicating some very low security apartments.
-Policy is needed to improve security for these apartments, and the observation that some scores are far higher than most are more thoroughly scrutinized later in the report.
+```{r figtwo,out.width="60%", fig.cap = "Histograms of the Distribution of Apartment Security levels", echo= FALSE}
+ggplot(rent_clean, aes(x=Security))+
+  geom_histogram(color = "black", fill = "white")+
+  theme_minimal()+
+  xlab("Level of security")
+```
 
-(Figure \@ref(fig:figtwo)), using ggplot2 (Wickham 2016), shows the average maintenance scores by ward.
-Cross referencing the income and demographics of each ward using (City of Toronto 2022) , we see that the higher income wards such as Toronto Saint Paul's and Toronto-Danforth (155,470 and 101,323 average yearly income respectively 72% white and 67% white respectively as of 2016) tend to have both higher means and medians than lower income wards, such as Davenport and Humber River-Black Creek (80,807 and 65,458 average yearly income as of 2016).
+(Figure \@ref(fig:figthree)) shows the average maintenance scores by ward.
+Cross referencing the income and demographics of each ward using (City of Toronto 2021), we see that the higher income wards such as Toronto Saint Paul's and Toronto-Danforth (155,470 and 101,323 average yearly income respectively as of 2016) tend to have both higher medians than lower income wards, such as Davenport and Humber River-Black Creek (80,807 and 65,458 average yearly income as of 2016).
 An exception seems to be Scarborough North which has a low yearly average income and high racial minority population but a high average apartment maintenance score.
 However note that the range is very small, implying that there are very few samples, so they may not be representative of the true state of apartments in this ward.
-Although there are several outlier Scores for each ward, in general apartment housing is better in richer wards than in poorer wards.
+Although there are some outliers, in general apartment housing is better in richer wards than in poorer wards.
 High income Wards also tend to have fewer racial minorities, with Toronto saint Paul and Toronto-Danforth having a 72% and 67% non-racial minority population respectively.
 
-```{r figtwo, fig.cap = "Boxplot of median apartment maintenance scores by ward", echo = TRUE}
+```{r figthree, out.width="60%", fig.cap = "Boxplot of Apartment Maintenance Scores by Ward", echo = FALSE}
 
@@
   coord_flip()+
-  labs(title = "Wealthier wards have better scores than poorer Wards",
-       x = "Ward",
-       y = "Apartment Score")
+  labs(x = "Ward",
+       y = "Maintenance Score")
 ```
 
-(Figure \@ref(fig:figthree)) shows the mean and median apartment scores by type.
-The results show that private buildings have the best average maintenance quality, and those from the Toronto housing Corporation have the lowest average.
-The spread of quality seems to be equal for all three housing types, so even accounting for variation private housing is consistently better than their public counterparts.
-This shows a disparity in priority for apartment maintenance and upkeep, as well as general housing inequality as those in social housing owned by the TCHC also deserve high average maintenance but do not, and reveals that the government are not on top of taking care of social housing.
-However it should be noted that (non TCHC owned) social housing maintenance is not very far behind private housing, showing a just upkeep of housing.
+(Figure \@ref(fig:figfour)) shows the median apartment scores by type.
+The results show that private apartment have the best  maintenance quality, and those from the Toronto Housing Corporation have the lowest average.
+The spread of quality (interquartile range) seems to be equal for all three housing types, so even accounting for variation private housing is consistently better than their public counterparts.
+However it should be noted that (non TCHC owned) social housing maintenance is not very far behind private housing, showing a just upkeep of housing by other institutions, so the lagging behind for TCHC housing shows lack of government upkeep or funding.
 Nevertheless according to (), people in social housing often have to take maintenance and upkeep into their own hands.
 
-```{r figthree, fig.cap = "Boxplot of median apartment scores by Apartment Type", echo = TRUE }
+```{r figfour, out.width="60%", fig.cap = "Boxplot of Apartment Maintenance Scores by Apartment Type", echo = FALSE }
 rent_clean %>%
@@
   geom_boxplot() +
-  labs(title = "Private Housing has better scores than social and TCHC housing ",
-       x = "Apartment Type",
-       y = "Apartment Score")
+  labs(x = "Apartment Type",
+       y = "Maintenance Score")
 ```
 
-(Table \@ref(tab:tabtwo))  shows the differences in security between each housing type.
-We see that overall the average secuirty scores are good for all types of apartment.
+(Table \@ref(tab:tabtwo)) shows the difference in security among each housing type.
+We see that overall the average security levels are good for all types of apartment.
 However the median shows that both TCHC owned and non TCHC owned social buildings have less stellar security than that of the private buildings, so their average scores were driven up by  a few very good apartments.
-This means that we can expect that a large number of social housing (which houses mainly low income populations) do not have the best security, which is counter-intuitive as poorer areas where this housing is often located are often in higher need of security due to correlation with crime rates.
+This means that we can expect that a large number of social housing (which houses mainly low income populations) do not have the best security, which is counter-intuitive as poorer areas where this type of housing is often located are often in higher need of more security due to higher correlation with crime rates.
 (cite poor housing live in crime areas).
-This could be alleviated with multi-income neighbourhoods, which places many different types of housing that can accomodate different income groups, but NIMBY-ism has slowed the progress of these neighbourhoods.
-```{r tabtwo}
+
+```{r tabtwo, echo = FALSE}
 rent_clean %>%
@@
   summarize(Mean = mean(Security),Median = median(Security)) %>%
-  knitr::kable(caption = "Table of Mean and Median security for each group") 
+  knitr::kable(caption = "Table of Mean and Median security for each Apartment Type")
 ```
 
-(Figure @\ref(fig:figfour)) shows the general trend of apartment maintenance scores for all types according to the year they were built.
+(Figure \@ref(fig:figfive)) shows the general trend of apartment maintenance scores for each apartment type according to the year they were built.
 Even though for all three types newer buildings have better maintenance Scores than their older counterparts, private housing still gets consistently higher scores than the other two types for almost all the years.
 Most private apartments were built between the 1960s to 1970s while social housing was built between the 1980s to 2000s, so even the newer age of the social housing buildings is not enough to protect against lower rates of maintenance.
-This shows either the persistence in lack of government and public maintenance or persistent bias in scoring, which both tie into income inequality.
+This shows either the persistent lack of government and public maintenance or persistent bias in scoring, which both tie into income inequality.
 It should be noted that the trend for private apartments may be heavily influenced by the points at the far right end, and that many private apartments also have low scores.
 We also removed an outlier, 1910 as it affected the accuracy of the trend for private apartments.
-```{r figfour,fig.cap = "Scatterplots and regression line of trend of apartment scores by type over 1950 - 2021", echo = FALSE }
+
+```{r figfive,out.width="60%", fig.cap = "Scatterplot and General Trend of Apartment Scores by Type from 1950 - 2021", echo = FALSE }
 rent_clean %>%
@@
   labs(title = "Privately owned apartment scores have ongiong advantage over social counterparts",
-       x = "Year",
-       y = "Score",
-       color = "Type") 
+       x = "Year Built",
+       y = "Maintenance Score",
+       color = "Apartment Type") 
 ```
 
-(Figure @\ref(fig:figfive)) shows that the wealthier wards have a larger percentage of their apartments scoring above 70 compared to lower income wards.
+(Figure \@ref(fig:figsix)) shows that wealthier wards have a larger percentage of their apartments scoring above 80 compared to lower income wards.
 Some of the poorer wards such as Humber River Black creek or Davenport do not have very highly rated apartments at all.
-The racial minority population in these two wards as of 2016 was 74% and 43% respectively (City of Toronto 2022).In 2016 the income for these wards was 65,548 and 80,807 dollars respectively while in the richer wards like Toronto Saint Paul and Etobicoke Centre it was 155,470 and 128,448 (City of Toronto 2022).
-Many other wards follow the same trend, with the wards with higher income having larger percentages of their apartments scoring above 70, with some lower income wards even having some percentage of their apartments have a score less than 50.
-This is another way to see the inequality in housing standards, and communicates that housing maintenance disproportionally caters to higher income areas.
-This is possibly because lower income people pay less rent, are less likely to protest (August 2016), or less general funding compared to previous decades (Tsenkova 2021).
+The racial minority population in these two wards as of 2016 was 74% and 43% respectively (City of Toronto 2021).
+In 2016 the income for these wards was 65,548 and 80,807 dollars respectively while in the richer wards like Toronto Saint Paul and Etobicoke Centre which both have a large percentage of highly rated apartments, it was 155,470 and 128,448 dollars respectively (City of Toronto 2021).
+Many other wards follow the same trend, with the wards with higher income having larger percentages of their apartments scoring above 80, while some lower income wards actually have a significant percentage of their apartments score less than 60.
+This is another way to see the inequality in housing standards, and communicates that housing maintenance disproportionately caters to higher income areas.
+This is possibly because lower income people pay less rent, are less likely to protest (August 2016), or because public housing receives less general funding compared to previous decades (Tsenkova 2021).
 note tsenkova states that less funding to build new housing
 
-```{r figfive,fig.cap = "Percentage of apartments in each ward with that have a score in a certain group", echo = FALSE }
+```{r figsix,out.width="60%", fig.cap = "Percentage of apartments in each ward with that have a score in a certain group", echo = FALSE }
 rent_clean %>%
-  mutate(Maintenance_Score = 
+  mutate(Score = 
            case_when(
@@
   coord_flip()+
-  geom_bar(mapping = aes(x = Ward, fill = Maintenance_Score),position="fill")+
-  scale_y_continuous(labels = scales::percent)
+  geom_bar(mapping = aes(x = Ward, fill = Score),position="fill")+
+  scale_y_continuous(labels = scales::percent)+
+  labs(x = "Ward",
+       y = " Percentage of Apartments in ward",
+       color = "Evaluation Score") 
 # Add Dark2 here 
 ```
-
-# 2.1 Results
 
@@
 
-As previously stated, the impact of these scores might lead to an inaccurate rating of social housing apartments or apartments in lower income areas as lower quality, thus perpetuating harmful views on low income people that further progress their marginalization in society by casting blame on social housing residents for poor conditions instead of the organizations supposed to keep maintenance (cite where poor are blamed for being dirty).
+As previously stated, the impact of these scores might lead to inaccurate judgments of social housing apartments or apartments in lower income areas as lower quality.
+This perpetuates harmful views on low income people that further progress their marginalization in society by casting blame on social housing residents for poor conditions instead of the organizations that are supposed to uphold maintenance (cite where poor are blamed for being dirty).
 Given the importance of these scores in rating apartments and housing choice, there is a need to understand the decisions that go into the scores and how they could be interpreted.
 
-It also brings into discussion the impact of housing on well being and the needs of the low income that have less than standard housing or require  lower income housing.
+This report brings into discussion the impact of housing on well being and the needs of the low income that have less than standard housing or require  lower income housing.
 It has been shown that public housing apartments have lower quality amenities than their private counterparts, and these can have several effects on the residents.
@@
 “incivilities” such as the presence of graffiti, boarded up houses, garbage accumulation...
-and broken windows, and the signalling effect of these physical features that no one cares; the perception of neighbourhoods as “safe” and under effective informal social controls...Others have postulated that neighbourhood conditions directly influence habits of privacy, child rearing, house keeping, and study, which can have an indirect effect on health."(Howden-Chapman 2004)
+and broken windows, and the signalling effect of these physical features that no one cares; the perception of neighbourhoods as “safe” and under effective informal social controls...Others have postulated that neighborhood conditions directly influence habits of privacy, child rearing, house keeping, and study, which can have an indirect effect on health."(Howden-Chapman 2004)
 
-There is also the need to discuss the shortcomings of the Toronto Housing Corporation and why they have not been up to date with maintenance of their apartments, as well as public policy that could be formed to improve the maintenance and security of social housing apartments in low income areas.
-Housing for the lower income is often substandard because they are at lower priority in almost every other facet as life such as law enforcement, schooling, etc.
-Either because they are often seen as net negative tax payers or that they do not have as much political persuasion (August 2016) as the middle class.
+There is also the need to discuss the shortcomings of the Toronto Housing Corporation and why they have not been up to date with maintenance of their apartments, as well as public policy that could be approved to improve the maintenance and security of social housing apartments in low income areas.
+Housing for the lower income is often substandard because they are at lower priority in almost every other facet as life such as law enforcement, schooling, etc., either because they are often seen as net negative tax payers or that they do not have as much political persuasion (August 2016) as the middle class.
 This intersects with a large number of poor being racial minorities, who are also often seen as less important in society.
@@
 
-Regarding social housing apartments that have less security, it is paradoxical as these people often live in more crime ridden areas that needs the extra security, bringing into question if their own lives are considered and the failings of public landlords and the TCHC to provide security akin to those of private apartment.
-Lower income people often do not have money to protect themselves like rich, so they rely on publicly provided assistance.
-(cite?) This also highlights lack of security for poor in other areas like biased law enforcement.
+The report has shown a disparity in priorities for apartment maintenance and upkeep between the public and private apartments, as well as general housing inequality as those in social housing owned by the TCHC do not have the high maintenance quality deserved.
+Regarding these apartments that have less security, it is paradoxical as these people often live in more crime ridden areas that needs the extra security, bringing into question if their own lives are considered and the failings of public landlords and the TCHC to provide security akin to those of private apartment.
+Lower income people often do not have money to protect themselves like the middle class and wealthy, so they rely on publicly provided assistance.
+(cite?) This can also be tied to the lack of security for the poor in other areas like law enforcement.
 (cite)
 
-The data showed that Scarborough North, which has a 90% visible minority population with an average income of 78,892 as of 2016, compared to the Toronto average of 100000 (cite toronto)) only has highly scored apartment buildings, an outlier.
-It is hopeful that average apartment scores for other wards with similar incomes increase in the future.
-However low income residents have concerns that a high level of security or good apartments in an otherwise lower income area can make the area more attractive to wealthier people, leading to gentrification, which pushes the original low income inhabitants out ((Evans, 2007: 6,7), Hulchanski,2006: Table – Change in Average Individual Income, 1970 to 2000, City of Toronto)(“It's all about power and you have none:” The marginalization of tenant resistance to mixed-income social housing redevelopment in Toronto, Canada,), which is a problem with the already tenuous social housing for these marginalized groups.(Hay,2005: 1)(cite actual)
+The data showed that Scarborough North, which has a 90% visible minority population with an average yearly income of 78,892 dollars as of 2016, compared to the Toronto average of 102,721 dollars (City of Toronto 2021), only has highly scored apartment buildings, an outlier.
+If this is not due to a sampling error, it is hopeful that apartment scores for other wards with similar incomes increase in the future.
+However low income residents have concerns that a high level of apartment security, or highly rated apartments in an otherwise lower income area can make the area more attractive to wealthier people.
+This could lead to gentrification, which pushes the original low income inhabitants out ((Evans, 2007: 6,7), Hulchanski,2006: Table – Change in Average Individual Income, 1970 to 2000, City of Toronto), which is a problem with the already tenuous social housing situation for these marginalized groups.(Hay,2005: 1)(cite actual)
 
-## Bias and ethical Concerns
+## Bias and Ethical Concerns
 
-There are some areas for possible bias in the data which perpetuate preconceived notions on marginalized groups.
-As  Rent Safe TO does not release its criteria for scoring, it is unknown what specific attributes they look for while grading.
-Measurement issues could occur as there is not a way to objectively measure an apartment (i.e it is unclear what distinguishes a rating of 4 from a rating of 5).
-It is thus possible to then give the an apartment attribute of the same quality a higher score in a wealthier ward/private building with other good amenities, whilst giving apartments in poorer wards/social housing a lower score due to bias against the general area or the social status of residents and not the apartment itself.
-This perpetuates biases against the lower income and minorities as people gets their previously held beliefs of the poor status of those communities confirmed (cite where people believe minority areas look worse).
-Lower scores are prevalent in poorer wards and wards with high visible minority populations, and biases against these people and the general "feel" of an area may influence people to rent a more expensive apartment with good scores when there's a cheaper one that was scored wrongly.
-On the other side, being in a more affluent ward may artificially prop up a mediocre aparmtent's score, thus inflating prices and making it less affordable even though its maintenance is lacking.
-There is thus a need for a more transparent grading system so that outsiders can cross check and make sure than no historic inequalities perpetuate themselves through scoring biased against poor or social housing folk.
+There are some areas for possible bias in the data which perpetuate preconceived notions of marginalized groups.
+As RentSafeTO does not release its criteria for grading, it is unknown what specific attributes they look for during the process.
+Measurement issues could occur as there is not a way to objectively grade an apartment i.e it is unclear what distinguishes a rating of 4 from a rating of 5.
+It is thus possible to then give an apartment attribute of the same quality a higher score in a wealthier ward/private building with other good amenities, whilst giving apartments in poorer wards/social housing a lower score, due to bias against the general area or the social status of residents and not because of the apartment block itself.
+This maintains biases against the lower income and minorities as people get their previously held beliefs of the poor status of those communities confirmed (cite where people believe minority areas look worse).
+Lower scores are prevalent in poorer wards and wards with high visible minority populations, and biases against these people and the general "feel" of an area may influence people to rent at a more expensive apartment block with good scores when there's a cheaper apartment that was inaccurately graded.
+On the other side, being in a more affluent ward may artificially prop up a mediocre apartment's score, thus inflating prices and making it less affordable even though its maintenance is lacking.
+There is thus a need for a more transparent grading system so that outsiders can cross check and make sure that no historic inequalities perpetuate themselves through biased scoring against those living or poor areas or social housing.
 
-## Collection problem
+## Data Collection Problem
 
-There are possible data collection problems with this data set.
-RentSafeTO collects data only for apartment blocks with 3 or more stories and ten or more units(cite), meaning that some buildings that are in dire need of maintenance may be left behind as the organization simply doesn't collect data for them, even though most people may see them as apartments (i.e a three story block of 9 flats).
-This leaves out a big piece of the full picture on apartment quality, and may also be why some known wealthier wards are getting higher average scores (the low income housing there does not count) and the effect of income inequality as poor people are more likely to live in such types of "apartments" than the wealthy (cite).
+There are some data collection problems with this data set.
+RentSafeTO collects data only for apartment blocks with 3 or more stories and ten or more units(cite), meaning that some buildings that are in dire need of maintenance may be left behind as the organization simply does not collect data for them, even though most people may see them as apartments (i.e a three story block of 9 flats).
+This leaves out a big piece of the full picture on apartment quality, and may also be why some known wealthier wards are getting higher average scores as the low income housing there does not count.
+It also leaves out the effect of income inequality as lower income people are more likely to live in such types of "apartments" than the wealthy (cite).
 For apartments that did qualify, the grading is a scheduled event organized with the landlord ahead of time (cite where this is stated on the website).
-This means that scores presented may not be entirely accurate for general upkeep of the building as landlords who neglect buildings can hurriedly make affairs before the evaluator/enforcement officer arrives.
-It may thus be necessary for graders to come up semi unannounced to get the true maintenance levels of the building for a more accurate scoring.
+This means that scores presented may not be an entirely accurate measure of the general upkeep of a building as landlords who neglect their property can hurriedly make affairs before the inspector/enforcement officer arrives.
+It may thus be necessary for inspectors to come up semi-unannounced to get the true evaluation of the building.
 
 ## Income
+
 It is important to note that not all living in private apartments are wealthy.
-In fact, there are many private apartments inhabited by low income people so this disparity in scores between private and social housing may be due to reasons other than income inequality.
+In fact, there are many private apartments inhabited mainly by low income people so this disparity in scores between private and social housing may be due to reasons other than income inequality.
+It is important not to overlook those who are struggling to pay rent to access these improved amenities.
+Not all those who live in private areas or wealthier wards are rich, as wealthier wards still have poor areas so their own experiences may be neglected.
 As stated earlier, less governmental priority to social apartments may be a major reason (Tsenkova 2021).
-The average income for each ward was an important variable as it would put other information into context, as it determines the priority in government spending not just in apartments but in general as lower income areas get less spending but wealthier income are often full of funding in housing, surroundings, and other outside amenities that could have gone into the scoring bias.
+
+The average income for each ward was an important variable because it would have put other information from the data into context, as it determines the priority in government funding not just in apartments but in general.
+This is because lower income areas get less funding but wealthier income areas are often full of funding when it comes to public infrastructure, outside maintenance, and other social amenities, all of which could have gone into the scoring bias.
 Unfortunately this variable was not included in the data set, so information on income had to come from a less up to date source.
 
-(should these be in next steps?)Private housing is now in itself getting harder to afford due to rising housing costs, forcing many so called middle class people into the need for social housing, or forcing them into lower rated apartments.
+(should these be in next steps?)Private housing is now in itself getting harder to afford due to rising housing costs, forcing many so called middle class people into social housing, or lower rated private apartments.
 It would be interesting to see the effect of rising prices and the movement of people that would have been considered high income into apartments in lower income wards or social housing, on total maintenance scores and security, to see if the government would then divert funding to these.
@@
 ## Weaknesses and next steps
-We should note that not everyone who lives in private housing is wealthy, and it is important not to overlook those who are struggling to pay rent to access those improved amenities.
-Not all those who live in private areas or wealthier wards are rich, wealth wards still have poor areas so their own experiences were neglected.
 
@@
 For next steps we could use the geo-location variable that was omitted for this report to see if the scores for apartments on boundaries between poorer or richer wards had the negative bias of the poor wards or positive bias of the rich.
-
-We could also compare the income and scores of buildings including those that did not have all amenities to further observe if there was inequality as apartments in poor areas tend to have complete absence of some amenities.
+We could also use this variable for a geo-spatial legend in contrast to the graphical legend used in the report.
+We could also include apartment blocks that did not have all amenities in the comparison of income and scores to further observe evidence of inequality as apartments in poor areas tend to have complete absence of some amenities.
 

### 2022-02-06 outputs/paper/paper.Rmd
Changed the text and added in text citations for rmd file, added more references to the bib file, changes some text in the R script, created a pdf from the rmd file.

@@
 abstract: "Housing maintenance benefits residents as it ensures proper functioning of ammenities, maintains structural integrity, and adds value to homes.
-However, previous research has found that priority in maintenance for public housing apartments and lower income apartments is little compared to their more privileged counterparts.We use data from Open data Toronto to further analyze this by using plots and tables to break down apartment scores and other attributes by Ward and apartment type,using R (R Core Team 2021) and other packages.
-We find that apartments in higher income areas and privately owned apartments tend to have better evaluation scores than those in lower income wards and public housing respectively.
+However, previous research has found that there is lower priority in maintenance for public housing apartments and lower income apartments compared to their more privileged counterparts.
+We use data from Open data Toronto to further analyze this by using plots and tables to break down apartment scores and other attributes by Ward and apartment type,using R [@citeR] and other packages.
+We find that apartments in higher income wards and privately owned apartments tend to have better evaluation scores than those in lower income wards and public housing respectively.
 This report shows the inequality in standards for housing in Toronto, as well as the need for better government accountability and housing justice.
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r global_options, include=FALSE}
+knitr::opts_chunk$set(fig.pos = 'H')
 library(tidyverse)
@@
 library(opendatatoronto)
+library(kableExtra)
 ```
@@
 Toronto is a large Canadian city divided into 25 wards.
-In order to make sure building maintenance standards are upheld and to allow apartment residents to look up safety scores of their apartments, Rent Safe TO has provides an apartment evaluation data set data set.
+In order to make sure building maintenance standards are upheld and to allow apartment residents to look up safety scores of their apartments, RentSafeTO provides an apartment evaluation data set data set.
 According to a set criteria not revealed to public, this data set rates apartment attributes according to a 5 point system, and then uses a formula to assign them an overall score over 100.
@@
 
-We obtained the data set from the City of Toronto Open Data Portal (https://open.toronto.ca/dataset/apartment-building-evaluation/) with `opendatatoronto` (Gelfand 2020), and cleaned and analyzed the data set using R (R Core Team 2021),tidyverse (Wickham et al.
-2019) and dplyr (@citetidyverse) packages.
-Figures and tables were created with ggplot2 (Wickham 2016) and kableextra (Zhu 2021).
-We created tables that first showed the data and important summary statistics, then went into showing the data for area and apartment types using box plots, linear model graphs and percentage bar charts, which highlighted inequality in housing evaluation scores.
+We obtained the data set from the City of Toronto Open Data Portal (https://open.toronto.ca/dataset/apartment-building-evaluation/) with `opendatatoronto` [@citeopentoronto], and cleaned and analyzed the data set using R [@citeR] and tidyverse [@citetidyverse] packages.
+Figures and tables were created with ggplot2 [@citeggplot], knitr [@citeknit] and kableExtra [@citekable].
+We first created tables and graphs that first showed the data.
+We then went into showing the summaries for area and apartment types using box plots, linear model graphs and percentage bar charts.
+These all highlighted inequality in housing evaluation scores.
 
-Not all apartments are equal, as this report shows that although general housing maintenance has increased in recent years (OECD 2020), there is still a large difference in quality of apartments around the city.
-Social housing, both operated by the TCHC and by other institutions, have lower average maintenance scores and lower median security scores than their private counterparts.
-Apartments in some wealthier wards also have a higher average maintenance score than those in poorer wards.
+Not all apartments are equal, as this report shows that although general housing maintenance has increased in recent years [@oecd], there is still a large difference in quality of apartments around the city.
+Social housing, both operated by the Toronto Housing Corporation (TCHC) and by other institutions, tend to have lower median and average maintenance scores than their private counterparts.
+Apartments in some wealthier wards also have a higher median maintenance score than those in poorer wards.
 Private apartments also tend to have better security than their social housing counterparts.
-According to (OECD 2020) and (Desroches and Poland, 2021), these disparities are due to a decline in public investment.
+According to [@oecd] and [@Desroches2021], these disparities are due to a decline in public investment.
 
-These scores are widely used to rate apartments ((Carter, 2004: 12)), however biased scoring and issues in measurement could create inaccurate property pricing, or otherwise give false impressions on the apartment block, and thus the type of people living there(Konadu-Agyemang et al.
-1994, pg 146).
-Low scores for public housing also exacerbate the already low view of people who need them (Konadu-Agyemang et al.
-1994, pg 146).
-This data and the statistics shows that vulnerable populations such as low income and minorities suffer systemic marginalization in facets of society, even in the housing sector as well maintained, secure housing is still unattainable for the city's most vulnerable, questioning the extent of the achievements of social housing.
-There are also concerns on the bias of the grading, and the impact of these scores on gentrification.
+These scores are widely used to rate apartments, however biased scoring and issues in measurement could create inaccurate property pricing, or otherwise give false impressions on the apartment block, and thus the type of people living there [@Konadu].
+Low scores for public housing also exacerbate the already low view of people who need them [@Konadu].
+This data and the statistics shows that vulnerable populations such as low income people and minorities suffer systemic marginalization in many facets of society.
+This includes the housing sector as well maintained, secure housing is still unattainable for the city's most vulnerable, questioning the extent of the achievements of social housing.
+There are also concerns on the bias of the grading, and the societal impact of these scores such as gentrification.
 
 # Data
-The data for this report was gotten from the Apartment Building Evaluation (Rent Safe TO 2022) from the Toronto Open Data Portal (city of Toronto 2022) using the R package opendatatoronto (Gelfand 2020).
+The data for this report was gotten from the Apartment Building Evaluation [@citedata] of the Toronto Open Data Portal (city of Toronto 2022) using the R package opendatatoronto [@citeopentoronto].
 This data is compiled by RentSafeTO and was last refreshed on January 22nd, 2022.
-Rent Safe TO collects data on apartments in the GTA that are 3 or more stories or had 10 more units (cite this), and response is mandatory for qualifying apartments, i.e the population.
-The raw data includes 9758 apartment blocks from the Greater Toronto Area and their categorical characteristics (40 variables in total) such as location and Type of property, as well as numeric characteristics such as Security,amenities and state of common areas.
-It also includes an overall score over 100 for each apartment block.
+Rent Safe TO collects data on apartments in the Greater Toronto Area (GTA) that are 3 or more stories tall or had 10 more units [@torontoeval], and response is mandatory for qualifying apartments, i.e the population.
+The raw data includes 9758 apartment blocks from the GTA and their categorical characteristics such as location and Type of property, as well as numeric characteristics such as Security,amenities and state of common areas.
+It also includes an overall score over 100 for each apartment block, and there are 40 variables in total.
 
@@
 That is, we wanted to compare apartments that a had poor/ lacking characteristic, not no characteristic at all.
+We thought it would also be helpful for the analysis to have the full picture in order to observe the possibility of bias in grading, instead of no grading at all.
 We removed all rows with missing values as they indicated one or more of the previous criteria was missing.
-It was also helpful to see the full picture to observe the possibility of bias in grading instead of no grading at all.
 The missing values seemed to be randomly dispersed about the data set so removing them would not cause much deviation.
@@
 We also included Security to observe the the direct relationship and differences in security between different groups of apartments.
-We then included property type, which states whether an apartment was privately owned, owned by the Toronto Housing Corporation (TCHC), or owned by some other non-private institution as our analysis focuses on social/affordable housing.
+We then included property type, which states whether an apartment was privately owned, owned by the TCHC, or owned by some other non-private institution, as our analysis focuses on social/affordable housing.
 The final data frame included 493 observations with 5 variables.
@@
 
-We are interested in the relationships between apartment Type and Ward on Score and Security, as well as the trend of Score over time.
+We are interested in the relationships between apartment Type, Ward,Score and Security, as well as the trend of Score over time.
 These relationships show a measure of income inequality in housing, and have possible implications for public policy.
-A preview of the data, (Table \@ref(tab:tabone)), (Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) using kableextra (Zhu 2021) and ggplot2 (Wickham 2016) respectively is shown below.
+The data, (Table \@ref(tab:tabone)), (Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) using kableextra (@citekable) and ggplot2 (@citeggplot) respectively is shown below.
 
-```{r tabone, fig.cap = "Table of the first few rows of the data", echo = FALSE}
+```{r tabone, echo = FALSE}
 
 x = head(rent_clean) %>%
-  knitr::kable(col.names = c("Score", "Ward", "Year built","Type", "Security"), caption = "Toronto Apartment Evaluation Data")
+  knitr::kable(col.names = c("Score", "Ward", "Year built","Type", "Security"), caption = "Toronto Apartment Evaluation Data") %>% 
+  kable_styling(latex_options = "HOLD_position")
 x
@@
 
-(Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) show the spread and distribution of apartments scores and their security levels.
-We see that the data for Scores and Age seems to be unimodal and normally distributed.
-There is an outlier of a very low apartment score, but it is not enough to skew the data to the left.
+(Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) show the spread and distribution of apartments scores and their security levels respectively.
+We see that the data for Scores seems to be unimodal and normally distributed.
+There is an outlier of a very low apartment score, but it is not enough to skew the data to the left.The median apartment score seems to be the same or a slightly above the mean apartment score, as there seems to be more apartments on the lower ends than there are on the higher end.
+This indicates there are many below average apartments.
 We can see from the figures that the average building has good maintenance and security, but is a bit old.
-The median apartment score seems to be the same or a slightly above the mean apartment score, as there seems to be more apartments on the lower ends than there are on the higher end.
-This indicates there are many below average apartments.
-As there are quite of lower security apartments, it seems that the mean Security will be lower than the median.
-some very low security apartments.
+As there are quite a lot of lower security apartments, it seems that the mean Security may be lower than the median.
 Policy is needed to improve security for these apartments, and the observation that some evaluation Scores are quite lower than most are more thoroughly scrutinized later in the report.
 
-```{r figone,out.width="60%", fig.cap = "Histogram of the Distribution of Apartment Scores", echo= FALSE}
-par(mfrow=c(2,2))
+```{r figone,out.width="70%", fig.cap = "Histogram of the Distribution of Apartment Scores", echo= FALSE}
+
 ggplot(rent_clean, aes(x=Score)) +
-  geom_histogram(color = "black", fill="white")+
+  geom_histogram(color = "black", fill="white",bins=30)+
   theme_minimal()+
@@
 
-```{r figtwo,out.width="60%", fig.cap = "Histograms of the Distribution of Apartment Security levels", echo= FALSE}
+```{r figtwo,out.width="70%", fig.cap = "Histogram of the Distribution of Apartment Security levels", echo= FALSE}
+
 ggplot(rent_clean, aes(x=Security))+
-  geom_histogram(color = "black", fill = "white")+
+  geom_histogram(color = "black", fill = "white",bins=30)+
   theme_minimal()+
@@
 (Figure \@ref(fig:figthree)) shows the average maintenance scores by ward.
-Cross referencing the income and demographics of each ward using (City of Toronto 2021), we see that the higher income wards such as Toronto Saint Paul's and Toronto-Danforth (155,470 and 101,323 average yearly income respectively as of 2016) tend to have both higher medians than lower income wards, such as Davenport and Humber River-Black Creek (80,807 and 65,458 average yearly income as of 2016).
+Cross referencing the income and demographics of each ward using [@torontoward], we see that the higher income wards such as Toronto Saint Paul  and Toronto-Danforth (155,470 and 101,323 average yearly income respectively as of 2016) tend to have higher median scores than lower income wards, such as Davenport and Humber River-Black Creek (80,807 and 65,458 average yearly income as of 2016).
 An exception seems to be Scarborough North which has a low yearly average income and high racial minority population but a high average apartment maintenance score.
-However note that the range is very small, implying that there are very few samples, so they may not be representative of the true state of apartments in this ward.
+However note that the range for this ward is very small, implying that there are very few samples, so they may not be representative of the true state of apartments in the ward.
 Although there are some outliers, in general apartment housing is better in richer wards than in poorer wards.
@@
 
-```{r figthree, out.width="60%", fig.cap = "Boxplot of Apartment Maintenance Scores by Ward", echo = FALSE}
+```{r figthree, out.width="70%", fig.cap = "Boxplot of Apartment Maintenance Scores by Ward", echo = FALSE}
 
@@
 (Figure \@ref(fig:figfour)) shows the median apartment scores by type.
-The results show that private apartment have the best  maintenance quality, and those from the Toronto Housing Corporation have the lowest average.
-The spread of quality (interquartile range) seems to be equal for all three housing types, so even accounting for variation private housing is consistently better than their public counterparts.
-However it should be noted that (non TCHC owned) social housing maintenance is not very far behind private housing, showing a just upkeep of housing by other institutions, so the lagging behind for TCHC housing shows lack of government upkeep or funding.
-Nevertheless according to (), people in social housing often have to take maintenance and upkeep into their own hands.
+The results show that private apartment have the best median maintenance quality, and those from the Toronto Housing Corporation have the lowest median.
+The range of quality (interquartile range) seems to be equal for all three housing types, so even accounting for variation private housing is consistently better than their public counterparts.
+However it should be noted that (non TCHC owned) social housing maintenance is not very far behind private housing, showing a just upkeep of housing by other institutions.
+This indicates that the disparity for TCHC housing comes from a lack of government upkeep or funding.
+Nevertheless people in social housing often have to take maintenance and upkeep into their own hands or otherwise enter legal battles [@Bryks2011].
 
-```{r figfour, out.width="60%", fig.cap = "Boxplot of Apartment Maintenance Scores by Apartment Type", echo = FALSE }
+```{r figfour, out.width="70%", fig.cap = "Boxplot of Apartment Maintenance Scores by Apartment Type", echo = FALSE }
 rent_clean %>%
@@
 However the median shows that both TCHC owned and non TCHC owned social buildings have less stellar security than that of the private buildings, so their average scores were driven up by  a few very good apartments.
-This means that we can expect that a large number of social housing (which houses mainly low income populations) do not have the best security, which is counter-intuitive as poorer areas where this type of housing is often located are often in higher need of more security due to higher correlation with crime rates.
-(cite poor housing live in crime areas).
+This means that we can expect that a large number of social housing (which houses mainly low income populations) do not have the best security, which is counter-intuitive as poorer areas where this type of housing is often located are often in higher need of more security due to higher correlation with crime rates [@Yen1999].
 
@@
   summarize(Mean = mean(Security),Median = median(Security)) %>%
-  knitr::kable(caption = "Table of Mean and Median security for each Apartment Type")
+  knitr::kable(caption = "Table of Mean and Median security for each Apartment Type")%>% 
+  kable_styling(latex_options = "HOLD_position")
 ```
@@
 It should be noted that the trend for private apartments may be heavily influenced by the points at the far right end, and that many private apartments also have low scores.
-We also removed an outlier, 1910 as it affected the accuracy of the trend for private apartments.
+We also removed an outlier, an apartment built in 1910 as it affected the accuracy of the trend for private apartments.
 
-```{r figfive,out.width="60%", fig.cap = "Scatterplot and General Trend of Apartment Scores by Type from 1950 - 2021", echo = FALSE }
+```{r figfive,out.width="70%", fig.cap = "Scatterplot and General Trend of Apartment Scores by Type from 1950 - 2021", echo = FALSE, message = FALSE }
 rent_clean %>%
@@
   theme_minimal() +
-  labs(title = "Privately owned apartment scores have ongiong advantage over social counterparts",
-       x = "Year Built",
+  labs(x = "Year Built",
        y = "Maintenance Score",
@@
 Some of the poorer wards such as Humber River Black creek or Davenport do not have very highly rated apartments at all.
-The racial minority population in these two wards as of 2016 was 74% and 43% respectively (City of Toronto 2021).
-In 2016 the income for these wards was 65,548 and 80,807 dollars respectively while in the richer wards like Toronto Saint Paul and Etobicoke Centre which both have a large percentage of highly rated apartments, it was 155,470 and 128,448 dollars respectively (City of Toronto 2021).
-Many other wards follow the same trend, with the wards with higher income having larger percentages of their apartments scoring above 80, while some lower income wards actually have a significant percentage of their apartments score less than 60.
+The racial minority population in these two wards as of 2016 was 74% and 43% respectively [@torontoward].
+In 2016 the income for these wards were 65,548 and 80,807 dollars respectively while in the richer wards like Toronto Saint Paul and Etobicoke Centre which both have a large percentage of highly rated apartments, they were 155,470 and 128,448 dollars respectively [@torontoward].
+Many other wards follow the same trend, with the wards with higher income having larger percentages of their apartments scoring above 80, while a few lower income wards actually have some percentage of their apartments score less than 60.
 This is another way to see the inequality in housing standards, and communicates that housing maintenance disproportionately caters to higher income areas.
-This is possibly because lower income people pay less rent, are less likely to protest (August 2016), or because public housing receives less general funding compared to previous decades (Tsenkova 2021).
-note tsenkova states that less funding to build new housing
+This is possibly because lower income people pay less rent, are less likely to protest [@August2016], or because public housing receives less general funding compared to previous decades [@Tsenkova2021].
 
-```{r figsix,out.width="60%", fig.cap = "Percentage of apartments in each ward with that have a score in a certain group", echo = FALSE }
+```{r figsix,out.width="70%", fig.cap = "Percentage of apartments in each ward with scores in a certain range", echo = FALSE }
 rent_clean %>%
@@
        color = "Evaluation Score") 
-# Add Dark2 here 
 ```
@@
 
-This report brings into discussion the impact of housing on well being and the needs of the low income that have less than standard housing or require  lower income housing.
+This report brings into discussion the impact of housing on well being and the needs of the low income that require low income housing or have less than standard housing.
 It has been shown that public housing apartments have lower quality amenities than their private counterparts, and these can have several effects on the residents.
@@
 “incivilities” such as the presence of graffiti, boarded up houses, garbage accumulation...
-and broken windows, and the signalling effect of these physical features that no one cares; the perception of neighbourhoods as “safe” and under effective informal social controls...Others have postulated that neighborhood conditions directly influence habits of privacy, child rearing, house keeping, and study, which can have an indirect effect on health."(Howden-Chapman 2004)
+and broken windows, and the signalling effect of these physical features that no one cares; the perception of neighbourhoods as “safe” and under effective informal social controls...Others have postulated that neighborhood conditions directly influence habits of privacy, child rearing, house keeping, and study, which can have an indirect effect on health." [@Chapman2004].
 
 There is also the need to discuss the shortcomings of the Toronto Housing Corporation and why they have not been up to date with maintenance of their apartments, as well as public policy that could be approved to improve the maintenance and security of social housing apartments in low income areas.
-Housing for the lower income is often substandard because they are at lower priority in almost every other facet as life such as law enforcement, schooling, etc., either because they are often seen as net negative tax payers or that they do not have as much political persuasion (August 2016) as the middle class.
-This intersects with a large number of poor being racial minorities, who are also often seen as less important in society.
-"Housing is merely one of many ways that systemic discrimination of poor folk and the lack of priority for them is shown.
-Lack of affordable or poor quality housing exacerbates challenges associated with low income, education, literacy, physical and mental disabilities, race, ethnicity and colour" (Carter, 2004: 25).
+Housing for the lower income is often substandard because they are at lower priority in many government operations [@Fleming2019] such as law enforcement, schooling, etc., either because they are often seen as net negative tax payers or that they are not as able to fight for their rights as the middle class [@Crosby2020].
+This intersects with a large number of low income people being racial minorities, who are also often seen as less important in society.
+Housing is one of the ways that governmental discrimination of poor people and the lack of priority for them is shown.
+Lack of affordable housing, low quality housing, and evictions increases problems associated with poverty, drug use, race, education, and physical and mental health [@Fleming2019].
 
-The report has shown a disparity in priorities for apartment maintenance and upkeep between the public and private apartments, as well as general housing inequality as those in social housing owned by the TCHC do not have the high maintenance quality deserved.
-Regarding these apartments that have less security, it is paradoxical as these people often live in more crime ridden areas that needs the extra security, bringing into question if their own lives are considered and the failings of public landlords and the TCHC to provide security akin to those of private apartment.
-Lower income people often do not have money to protect themselves like the middle class and wealthy, so they rely on publicly provided assistance.
-(cite?) This can also be tied to the lack of security for the poor in other areas like law enforcement.
+The report has shown a disparity in priorities for apartment maintenance and upkeep between public and private apartments, as well as general housing inequality as those in social housing owned by the TCHC do not have the high apartment quality or security they deserve.
+Regarding these apartments that have less security, it is paradoxical as these people often live in more crime ridden areas that require the extra security [@Stretesky2004], bringing into question if their own lives are considered, as well a the failings of public landlords and the TCHC to provide security akin to those of private apartments.
+Lower income people often do not have money to protect themselves like the middle class and wealthy, so they rely on publicly provided assistance (cite?) This can also be tied to the lack of non-violent security for the poor in other areas like law enforcement.
 (cite)
 
-The data showed that Scarborough North, which has a 90% visible minority population with an average yearly income of 78,892 dollars as of 2016, compared to the Toronto average of 102,721 dollars (City of Toronto 2021), only has highly scored apartment buildings, an outlier.
+The data showed that Scarborough North, which has a 90% visible minority population with an average yearly income of 78,892 dollars as of 2016 compared to the Toronto average of 102,721 dollars [@torontoward], only has highly scored apartment buildings.
 If this is not due to a sampling error, it is hopeful that apartment scores for other wards with similar incomes increase in the future.
 However low income residents have concerns that a high level of apartment security, or highly rated apartments in an otherwise lower income area can make the area more attractive to wealthier people.
-This could lead to gentrification, which pushes the original low income inhabitants out ((Evans, 2007: 6,7), Hulchanski,2006: Table – Change in Average Individual Income, 1970 to 2000, City of Toronto), which is a problem with the already tenuous social housing situation for these marginalized groups.(Hay,2005: 1)(cite actual)
+This could lead to gentrification, which pushes the original low income inhabitants out, which is a problem with the already tenuous social housing situation for these marginalized groups [@August2014] and [@Crosby2020].
 
@@
 As RentSafeTO does not release its criteria for grading, it is unknown what specific attributes they look for during the process.
-Measurement issues could occur as there is not a way to objectively grade an apartment i.e it is unclear what distinguishes a rating of 4 from a rating of 5.
+Measurement issues could occur as there is not a way to objectively grade an apartment, i.e it is unclear what distinguishes a rating of 4 from a rating of 5.
 It is thus possible to then give an apartment attribute of the same quality a higher score in a wealthier ward/private building with other good amenities, whilst giving apartments in poorer wards/social housing a lower score, due to bias against the general area or the social status of residents and not because of the apartment block itself.
-This maintains biases against the lower income and minorities as people get their previously held beliefs of the poor status of those communities confirmed (cite where people believe minority areas look worse).
+This maintains biases against the lower income and minorities as people get their previously held beliefs about the poor status of those communities confirmed [@Kamphuis2010].
 Lower scores are prevalent in poorer wards and wards with high visible minority populations, and biases against these people and the general "feel" of an area may influence people to rent at a more expensive apartment block with good scores when there's a cheaper apartment that was inaccurately graded.
@@
 There are some data collection problems with this data set.
-RentSafeTO collects data only for apartment blocks with 3 or more stories and ten or more units(cite), meaning that some buildings that are in dire need of maintenance may be left behind as the organization simply does not collect data for them, even though most people may see them as apartments (i.e a three story block of 9 flats).
+RentSafeTO collects data only for apartment blocks with 3 or more stories or ten or more units[@torontoeval], meaning that some buildings that are in dire need of maintenance may be left behind as the organization simply does not collect data for them, even though most people may see them as apartments (i.e a three story block of 9 flats).
 This leaves out a big piece of the full picture on apartment quality, and may also be why some known wealthier wards are getting higher average scores as the low income housing there does not count.
-It also leaves out the effect of income inequality as lower income people are more likely to live in such types of "apartments" than the wealthy (cite).
-For apartments that did qualify, the grading is a scheduled event organized with the landlord ahead of time (cite where this is stated on the website).
+It also leaves out the effect of income inequality as lower income people are now more likely to live in such types of "apartments" than the wealthy [@Musterd2014].
+For apartments that did qualify, the grading is a scheduled event organized with the landlord ahead of time [@torontoeval].
 This means that scores presented may not be an entirely accurate measure of the general upkeep of a building as landlords who neglect their property can hurriedly make affairs before the inspector/enforcement officer arrives.
@@
 Not all those who live in private areas or wealthier wards are rich, as wealthier wards still have poor areas so their own experiences may be neglected.
-As stated earlier, less governmental priority to social apartments may be a major reason (Tsenkova 2021).
 
-The average income for each ward was an important variable because it would have put other information from the data into context, as it determines the priority in government funding not just in apartments but in general.
+The average income for each ward was an important variable because it would have put other information from the data into context, as it determines the priority in government funding not just for apartments but in general.
 This is because lower income areas get less funding but wealthier income areas are often full of funding when it comes to public infrastructure, outside maintenance, and other social amenities, all of which could have gone into the scoring bias.
@@
 While the N/As were dispersed randomly throughout the data set, removing them may have caused bias in the data set so our means, medians and analysis of Scores may have been different if the N/A values had been filled.
-Also, removing variables even if they were already present in the Score variable may have caused incorrect analysis as there may be some important information in the rejected variables that either brings a new perspective or aids one of the accepted variable in explaining the relationships between apartment type, ward, and maintenance scores.
+Also, removing variables even if they were already present in the Score variable may have caused incorrect analysis as there may be some important information in the rejected variables that either brings a new perspective or aids one of the accepted variable in explaining the relationships between apartment type, ward, maintenance scores and security.
 
@@
 https://link.springer.com/chapter/10.1007/978-3-030-69563-7_2
+
+<div id="refs"></div>

### 2022-02-06 scripts/01-data_cleaning.R

@@ -6,16 +6,13 @@
 # License: MIT
 # Pre-requisites: 
 # - Need to have installed the opendatatoronto and tidyr packages. 
-# - Gitignore it!
-# - Change these to yours
-# Any other information needed?
+
 
 
 #### Workspace setup ####
-# Use R Projects, not setwd().
+
 library(opendatatoronto)
 library(tidyverse)
-
 # Read in the raw data.
 apartment_data = list_package_resources("4ef82789-e038-44ef-a478-a8f3590c3eb1") %>%
   filter(name=="Apartment Building Evaluation") %>%

### 2022-02-06 outputs/paper/paper.Rmd
Changes some text, created a pdf based on the changes.

@@
 We find that apartments in higher income wards and privately owned apartments tend to have better evaluation scores than those in lower income wards and public housing respectively.
-This report shows the inequality in standards for housing in Toronto, as well as the need for better government accountability and housing justice.
-Improving the gap in aparment quality helps improve the lives of poor and marginalized people."
+This report shows the inequality in standards for housing in Toronto, as well as the need for better government accountability and housing justice, as improving the gap in aparment quality helps improve the lives of poor and marginalized people."
 output:
@@
 # Data
+
+## Data Source
+
 The data for this report was gotten from the Apartment Building Evaluation [@citedata] of the Toronto Open Data Portal (city of Toronto 2022) using the R package opendatatoronto [@citeopentoronto].
@@
 
+## Data Collection and cleaning
 For data cleaning, our intended sample was apartments that included all available rooms, amenities, and other characteristics.
@@
 
+## Data Visualization
+
 ```{r,include=FALSE}
@@
 These relationships show a measure of income inequality in housing, and have possible implications for public policy.
-The data, (Table \@ref(tab:tabone)), (Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) using kableextra (@citekable) and ggplot2 (@citeggplot) respectively is shown below.
+The data, (Table \@ref(tab:tabone)), (Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) using kableextra (@citekable), knitr [@citeknit] and ggplot2 (@citeggplot) respectively are shown below.
 
@@
 
-```{r figone,out.width="70%", fig.cap = "Histogram of the Distribution of Apartment Scores", echo= FALSE}
+```{r figone,out.width="90%", fig.cap = "Histogram of the Distribution of Apartment Scores", echo= FALSE}
 
@@
 
-```{r figtwo,out.width="70%", fig.cap = "Histogram of the Distribution of Apartment Security levels", echo= FALSE}
+```{r figtwo,out.width="90%", fig.cap = "Histogram of the Distribution of Apartment Security levels", echo= FALSE}
 
@@
 
-```{r figthree, out.width="70%", fig.cap = "Boxplot of Apartment Maintenance Scores by Ward", echo = FALSE}
+```{r figthree, out.width="90%", fig.cap = "Boxplot of Apartment Maintenance Scores by Ward", echo = FALSE}
 
@@
 
-```{r figfour, out.width="70%", fig.cap = "Boxplot of Apartment Maintenance Scores by Apartment Type", echo = FALSE }
+```{r figfour, out.width="90%", fig.cap = "Boxplot of Apartment Maintenance Scores by Apartment Type", echo = FALSE }
 rent_clean %>%
@@
 
-```{r figfive,out.width="70%", fig.cap = "Scatterplot and General Trend of Apartment Scores by Type from 1950 - 2021", echo = FALSE, message = FALSE }
+```{r figfive,out.width="90%", fig.cap = "Scatterplot and General Trend of Apartment Scores by Type from 1950 - 2021", echo = FALSE, message = FALSE }
 rent_clean %>%
@@
   theme_minimal() +
+  scale_color_manual(values = c("PRIVATE" = "brown", "SOCIAL HOUSING" = "black", "TCHC" = "darkorange")) +
   labs(x = "Year Built",
@@
 
-```{r figsix,out.width="70%", fig.cap = "Percentage of apartments in each ward with scores in a certain range", echo = FALSE }
+```{r figsix,out.width="90%", fig.cap = "Percentage of apartments in each ward with scores in a certain range", echo = FALSE }
 rent_clean %>%
@@
   scale_y_continuous(labels = scales::percent)+
+  scale_color_manual("YlOrRd")+
   labs(x = "Ward",
@@
 
-# Discussion
+## Data discussion
 
-## Societal Impact of these scores
+### Societal Impact of Scores
 
@@
 
-## Bias and Ethical Concerns
+### Bias and Ethical Concerns
 
@@
 
-## Data Collection Problem
+### Data Collection Problems
 
@@
 
-## Income
+### Income
 
@@
 # References
-1.  OECD (2020), “Social housing: A key part of past and future housing policy”, Employment, Labour and Social Affairs Policy Briefs, OECD, Paris, http://oe.cd/social-housing-2020
-
-It's all about power and you have none:” The marginalization of tenant resistance to mixed-income social housing redevelopment in Toronto, Canada
-
-https://www.toronto.ca/city-government/data-research-maps/neighbourhoods-communities/ward-profiles/
-
-Howden-Chapman PHousing standards: a glossary of housing and healthJournal of Epidemiology & Community Health 2004;58:162-168.
-
-https://link.springer.com/chapter/10.1007/978-3-030-69563-7_2
-
-<div id="refs"></div>

### 2022-02-06 outputs/paper/paper.Rmd
Finishing touches in writing and citations.

@@
 
-We obtained the data set from the City of Toronto Open Data Portal (https://open.toronto.ca/dataset/apartment-building-evaluation/) with `opendatatoronto` [@citeopentoronto], and cleaned and analyzed the data set using R [@citeR] and tidyverse [@citetidyverse] packages.
+We obtained the data set from the City of Toronto Open Data Portal (https://open.toronto.ca/dataset/apartment-building-evaluation/) with `opendatatoronto` [@citeopentoronto], and cleaned and analyzed the data set using R [@citeR], dplyr, [@citedplyr] and tidyverse [@citetidyverse] packages.
 Figures and tables were created with ggplot2 [@citeggplot], knitr [@citeknit] and kableExtra [@citekable].
@@
 
-## Data Source
+## Data Source and collection
 
-The data for this report was gotten from the Apartment Building Evaluation [@citedata] of the Toronto Open Data Portal (city of Toronto 2022) using the R package opendatatoronto [@citeopentoronto].
-This data is compiled by RentSafeTO and was last refreshed on January 22nd, 2022.
+The data for this report was gotten from the Apartment Building Evaluation [@citedata] of the Toronto Open Data Portal [@citewebsite] using the R package opendatatoronto [@citeopentoronto].
+This data is compiled by RentSafeTO and was last accessed on January 22nd, 2022.
 Rent Safe TO collects data on apartments in the Greater Toronto Area (GTA) that are 3 or more stories tall or had 10 more units [@torontoeval], and response is mandatory for qualifying apartments, i.e the population.
@@
 
-## Data Collection and cleaning
+## Data cleaning
 For data cleaning, our intended sample was apartments that included all available rooms, amenities, and other characteristics.
@@
 
-The report has shown a disparity in priorities for apartment maintenance and upkeep between public and private apartments, as well as general housing inequality as those in social housing owned by the TCHC do not have the high apartment quality or security they deserve.
+The report has shown a disparity in priorities for apartment maintenance and upkeep between public and private apartments, as well as general housing inequality as those in social housing owned by the TCHC do not have the high apartment quality or security they deserve [@rights].
 Regarding these apartments that have less security, it is paradoxical as these people often live in more crime ridden areas that require the extra security [@Stretesky2004], bringing into question if their own lives are considered, as well a the failings of public landlords and the TCHC to provide security akin to those of private apartments.
-Lower income people often do not have money to protect themselves like the middle class and wealthy, so they rely on publicly provided assistance (cite?) This can also be tied to the lack of non-violent security for the poor in other areas like law enforcement.
-(cite)
+Lower income people often do not have money to protect themselves like the middle class and wealthy, so they rely on publicly provided assistance (cite?)
 
@@
 (should these be in next steps?)Private housing is now in itself getting harder to afford due to rising housing costs, forcing many so called middle class people into social housing, or lower rated private apartments.
-It would be interesting to see the effect of rising prices and the movement of people that would have been considered high income into apartments in lower income wards or social housing, on total maintenance scores and security, to see if the government would then divert funding to these.
+It would be interesting to see the effect of rising prices and the movement of people that would have been considered high income into apartments in lower income wards or social housing, on total maintenance scores and security, thus causing a hogher demand for those housing and tighter stretching available funds [@irwin].
 

### 2022-02-07 outputs/paper/paper.Rmd
Final grammar corrections.

@@
 We find that apartments in higher income wards and privately owned apartments tend to have better evaluation scores than those in lower income wards and public housing respectively.
+Private apartments also have higher levels of security than social housing.
 This report shows the inequality in standards for housing in Toronto, as well as the need for better government accountability and housing justice, as improving the gap in aparment quality helps improve the lives of poor and marginalized people."
@@
 According to a set criteria not revealed to public, this data set rates apartment attributes according to a 5 point system, and then uses a formula to assign them an overall score over 100.
-In this report, we are interested in how attributes such as relative wealth of the ward and the type of apartment affects the total maintenance score of an apartment, and how these scores have changed over the years.
+In this report, we are interested in how attributes such as relative wealth of the ward and the type of apartment affects the total maintenance score of an apartment and its level of security, as well as how these scores have changed over the years.
 
-We obtained the data set from the City of Toronto Open Data Portal (https://open.toronto.ca/dataset/apartment-building-evaluation/) with `opendatatoronto` [@citeopentoronto], and cleaned and analyzed the data set using R [@citeR], dplyr, [@citedplyr] and tidyverse [@citetidyverse] packages.
-Figures and tables were created with ggplot2 [@citeggplot], knitr [@citeknit] and kableExtra [@citekable].
-We first created tables and graphs that first showed the data.
-We then went into showing the summaries for area and apartment types using box plots, linear model graphs and percentage bar charts.
+We obtained the data set from the City of Toronto Open Data Portal [@citewebsite] with `opendatatoronto` [@citeopentoronto], and cleaned and analyzed the data set using `R` [@citeR], `dplyr`, [@citedplyr] and `tidyverse` [@citetidyverse] packages.
+Figures and tables were created with `ggplot2` [@citeggplot], `knitr` [@citeknit] and `kableExtra` [@citekable].
+We first created tables and graphs that showed the data.
+We then went into showing the relevant information for apartment types and wards using box plots, linear model graphs and percentage bar charts.
 These all highlighted inequality in housing evaluation scores.
@@
 Not all apartments are equal, as this report shows that although general housing maintenance has increased in recent years [@oecd], there is still a large difference in quality of apartments around the city.
-Social housing, both operated by the Toronto Housing Corporation (TCHC) and by other institutions, tend to have lower median and average maintenance scores than their private counterparts.
+Social housing, both operated by the Toronto Housing Corporation (TCHC) and by other institutions, tend to have lower median maintenance scores than their private counterparts.
+These higher scores for private apartments have been consistent over the years.
 Apartments in some wealthier wards also have a higher median maintenance score than those in poorer wards.
+These wealthier wards also have a higher amount of high quality apartment blocks.
 Private apartments also tend to have better security than their social housing counterparts.
-According to [@oecd] and [@Desroches2021], these disparities are due to a decline in public investment.
+According to [@oecd] and [@Desroches2021], these disparities are due to a decline in public investment in social housing.
 
@@
 This includes the housing sector as well maintained, secure housing is still unattainable for the city's most vulnerable, questioning the extent of the achievements of social housing.
-There are also concerns on the bias of the grading, and the societal impact of these scores such as gentrification.
+There are also concerns about the possible grading bias, as well as the societal impact of these scores such as gentrification.
 
@@
 Rent Safe TO collects data on apartments in the Greater Toronto Area (GTA) that are 3 or more stories tall or had 10 more units [@torontoeval], and response is mandatory for qualifying apartments, i.e the population.
+RentSafeTO collects the data by setting up an appointment with the landlord of the apartment block for inspection, and take notes and photographs of their inspection to upload to the City’s mobile investigation application.
+Later, the total building evaluation score is calculated and made available to review and post on the Tenant Notification Board [@torontoeval].
 The raw data includes 9758 apartment blocks from the GTA and their categorical characteristics such as location and Type of property, as well as numeric characteristics such as Security,amenities and state of common areas.
@@
 This was because we wanted an "all things being equal" analysis, the assumption that amenities wise, the apartments were alike.
-That is, we wanted to compare apartments that a had poor/ lacking characteristic, not no characteristic at all.
-We thought it would also be helpful for the analysis to have the full picture in order to observe the possibility of bias in grading, instead of no grading at all.
+That is, we wanted to compare apartments that had poor/lacking characteristics, not no characteristic at all.
+We thought it would also be helpful for the analysis to have the full picture in order to observe the possibility of bias in grading.
 We removed all rows with missing values as they indicated one or more of the previous criteria was missing.
@@
 
-We included a variable if there was anything important that made it stand out for our analysis.
+We included a variable in our data frame if there was anything important that made it stand out for our analysis.
 We omitted most variables in the original set as their direct effects were not important to the analysis, and the total score variable encompassed all of their individual scores.
 we first removed _id and RSN as they were merely ID variables.
-We also excluded the ward number variable as the ward names variable (Ward) was easier to use for identification.
+We also excluded the numeric ward variable as the ward names variable, which contains the 25 Toronto wards, was easier to use for identification.
 We included year built as we wanted to see the changes in apartment evaluation between older and newer apartments.
 We included score as it was the main determinant of maintenance standards.
-We also included Security to observe the the direct relationship and differences in security between different groups of apartments.
+We also included Security (which measures the condition of security systems of the apartment) to observe the the direct relationship and differences in security between different groups of apartments.
 We then included property type, which states whether an apartment was privately owned, owned by the TCHC, or owned by some other non-private institution, as our analysis focuses on social/affordable housing.
@@
 ```{r,include=FALSE}
+# loading in data from r script
 rent_clean = read_csv(here::here("inputs/data/rent_clean.csv"))
@@
 These relationships show a measure of income inequality in housing, and have possible implications for public policy.
-The data, (Table \@ref(tab:tabone)), (Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) using kableextra (@citekable), knitr [@citeknit] and ggplot2 (@citeggplot) respectively are shown below.
+The data, (Table \@ref(tab:tabone)), (Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) using `kableextra` [@citekable], `knitr` [@citeknit] and `ggplot2` [@citeggplot] respectively are shown below.
 
 ```{r tabone, echo = FALSE}
-
+# creating table of main variables
 x = head(rent_clean) %>%
   knitr::kable(col.names = c("Score", "Ward", "Year built","Type", "Security"), caption = "Toronto Apartment Evaluation Data") %>% 
-  kable_styling(latex_options = "HOLD_position")
+  kable_styling(latex_options = "HOLD_position") # keeping the table in place
 x
@@
 We see that the data for Scores seems to be unimodal and normally distributed.
-There is an outlier of a very low apartment score, but it is not enough to skew the data to the left.The median apartment score seems to be the same or a slightly above the mean apartment score, as there seems to be more apartments on the lower ends than there are on the higher end.
-This indicates there are many below average apartments.
-We can see from the figures that the average building has good maintenance and security, but is a bit old.
-As there are quite a lot of lower security apartments, it seems that the mean Security may be lower than the median.
+There is an outlier of a very low apartment score, but it is not enough to skew the data to the left.The median apartment score seems to be the same or a slightly above the mean apartment score, as there seems to be more apartments on the lower end than there are on the higher end.
+This indicates that there are many below average apartments.
+We can see from the figures that the average building has good maintenance and a high standard of security.
+However as there are still quite a lot of lower security apartments, it seems that the mean level of security may be lower than the median.
 Policy is needed to improve security for these apartments, and the observation that some evaluation Scores are quite lower than most are more thoroughly scrutinized later in the report.
@@
 ggplot(rent_clean, aes(x=Score)) +
-  geom_histogram(color = "black", fill="white",bins=30)+
+  geom_histogram(color = "black", fill="white",bins=30)+ #to remove the bins=30 error
   theme_minimal()+
-  xlab("Evaluation Score") 
+  xlab("Evaluation Score") +
+  ylab("Number of Apartments")
 ```
@@
   theme_minimal()+
-  xlab("Level of security") 
+  xlab("Level of security")+
+  ylab("Number of Apartments")
 ```
 
-(Figure \@ref(fig:figthree)) shows the average maintenance scores by ward.
-Cross referencing the income and demographics of each ward using [@torontoward], we see that the higher income wards such as Toronto Saint Paul  and Toronto-Danforth (155,470 and 101,323 average yearly income respectively as of 2016) tend to have higher median scores than lower income wards, such as Davenport and Humber River-Black Creek (80,807 and 65,458 average yearly income as of 2016).
-An exception seems to be Scarborough North which has a low yearly average income and high racial minority population but a high average apartment maintenance score.
+(Figure \@ref(fig:figthree)) shows the median and range of maintenance scores by ward.
+Cross referencing the income and demographics of each ward using [@torontoward], we see that the higher income wards such as Toronto Saint Paul  and Toronto-Danforth (155,470 and 101,323 dollars for average yearly income respectively as of 2016) tend to have higher median scores than lower income wards, such as Davenport and Humber River-Black Creek (80,807 and 65,458 dollars for average yearly income as of 2016).
+An exception seems to be Scarborough North which has a low yearly average income and high racial minority population, but a high average apartment maintenance score.
 However note that the range for this ward is very small, implying that there are very few samples, so they may not be representative of the true state of apartments in the ward.
-Although there are some outliers, in general apartment housing is better in richer wards than in poorer wards.
+Although there are some outliers, in general apartments are better in richer wards than in poorer wards.
 High income Wards also tend to have fewer racial minorities, with Toronto saint Paul and Toronto-Danforth having a 72% and 67% non-racial minority population respectively.
@@
   geom_boxplot() +
-  coord_flip()+
+  coord_flip()+ # flipping the graph to fit all the ward names
   labs(x = "Ward",
@@
 (Figure \@ref(fig:figfour)) shows the median apartment scores by type.
-The results show that private apartment have the best median maintenance quality, and those from the Toronto Housing Corporation have the lowest median.
+The results show that private apartments have the best median maintenance quality, and those from the Toronto Housing Corporation have the lowest.
 The range of quality (interquartile range) seems to be equal for all three housing types, so even accounting for variation private housing is consistently better than their public counterparts.
@@
 This indicates that the disparity for TCHC housing comes from a lack of government upkeep or funding.
-Nevertheless people in social housing often have to take maintenance and upkeep into their own hands or otherwise enter legal battles [@Bryks2011].
+Nevertheless people in social housing often have to take maintenance and upkeep into their own hands or otherwise enter legal battles, as shown by [@Bryks2011].
 
 ```{r figfour, out.width="90%", fig.cap = "Boxplot of Apartment Maintenance Scores by Apartment Type", echo = FALSE }
+
 rent_clean %>%
@@
 (Figure \@ref(fig:figfive)) shows the general trend of apartment maintenance scores for each apartment type according to the year they were built.
-Even though for all three types newer buildings have better maintenance Scores than their older counterparts, private housing still gets consistently higher scores than the other two types for almost all the years.
+We see that most apartments in the sample are quite old, being built between the 1950s to 1980s.
+Even though for all three types of apartment, newer buildings have better maintenance Scores than their older counterparts, private housing still gets consistently higher scores than the other two types for almost all the years.
 Most private apartments were built between the 1960s to 1970s while social housing was built between the 1980s to 2000s, so even the newer age of the social housing buildings is not enough to protect against lower rates of maintenance.
@@
   geom_point() +
-  geom_smooth(method = 'lm', se = F)+
-  geom_abline()+
+  geom_smooth(method = 'lm', se = F)+ # to see the general trend
+  geom_abline()+ # The regression line
   theme_minimal() +
-  scale_color_manual(values = c("PRIVATE" = "brown", "SOCIAL HOUSING" = "black", "TCHC" = "darkorange")) +
+  scale_color_manual(values = c("PRIVATE" = "brown", "SOCIAL HOUSING" = "black", "TCHC" = "darkorange")) + # custom colors
   labs(x = "Year Built",
@@
 The racial minority population in these two wards as of 2016 was 74% and 43% respectively [@torontoward].
-In 2016 the income for these wards were 65,548 and 80,807 dollars respectively while in the richer wards like Toronto Saint Paul and Etobicoke Centre which both have a large percentage of highly rated apartments, they were 155,470 and 128,448 dollars respectively [@torontoward].
+In 2016 the average annual income for these wards were 65,548 and 80,807 dollars respectively while in the richer wards like Toronto Saint Paul and Etobicoke Centre which both have a large percentage of highly rated apartments, the yearly average incomes were 155,470 and 128,448 dollars respectively [@torontoward].
 Many other wards follow the same trend, with the wards with higher income having larger percentages of their apartments scoring above 80, while a few lower income wards actually have some percentage of their apartments score less than 60.
@@
              TRUE ~ "None"
-             )
+             ) # group scores into separate categories
   ) %>%
@@
   coord_flip()+
-  geom_bar(mapping = aes(x = Ward, fill = Score),position="fill")+
-  scale_y_continuous(labels = scales::percent)+
+  geom_bar(mapping = aes(x = Ward, fill = Score),position="fill")+ # coloring the different categories
+  scale_y_continuous(labels = scales::percent)+ # getting the percentage of scores within each ward
   scale_color_manual("YlOrRd")+
@@
 As previously stated, the impact of these scores might lead to inaccurate judgments of social housing apartments or apartments in lower income areas as lower quality.
-This perpetuates harmful views on low income people that further progress their marginalization in society by casting blame on social housing residents for poor conditions instead of the organizations that are supposed to uphold maintenance (cite where poor are blamed for being dirty).
+This perpetuates harmful views on low income people that further progresses their marginalization in society by casting blame on social housing residents for poor conditions instead of the organizations that are supposed to maintain apartment quality.
 Given the importance of these scores in rating apartments and housing choice, there is a need to understand the decisions that go into the scores and how they could be interpreted.
@@
 The report has shown a disparity in priorities for apartment maintenance and upkeep between public and private apartments, as well as general housing inequality as those in social housing owned by the TCHC do not have the high apartment quality or security they deserve [@rights].
-Regarding these apartments that have less security, it is paradoxical as these people often live in more crime ridden areas that require the extra security [@Stretesky2004], bringing into question if their own lives are considered, as well a the failings of public landlords and the TCHC to provide security akin to those of private apartments.
-Lower income people often do not have money to protect themselves like the middle class and wealthy, so they rely on publicly provided assistance (cite?)
+Regarding these apartments that have less security, it is paradoxical as these people often live in more crime ridden areas that require the extra security [@Stretesky2004].
+This brings into question if their own lives are considered, as well as the failings of public landlords and the TCHC to provide security akin to those of private apartments.
+Lower income people often do not have money to protect themselves like the wealthy, so they need to rely on publicly provided assistance.
 
@@
 There are some areas for possible bias in the data which perpetuate preconceived notions of marginalized groups.
-As RentSafeTO does not release its criteria for grading, it is unknown what specific attributes they look for during the process.
+As RentSafeTO does not release its criteria for grading, it is unknown what specific qualities they look for during the process.
 Measurement issues could occur as there is not a way to objectively grade an apartment, i.e it is unclear what distinguishes a rating of 4 from a rating of 5.
@@
 On the other side, being in a more affluent ward may artificially prop up a mediocre apartment's score, thus inflating prices and making it less affordable even though its maintenance is lacking.
-There is thus a need for a more transparent grading system so that outsiders can cross check and make sure that no historic inequalities perpetuate themselves through biased scoring against those living or poor areas or social housing.
+There is thus a need for a more transparent grading system so that outsiders can cross check and make sure that no historic inequalities perpetuate themselves through biased scoring against those living in poor areas or social housing.
 
-### Data Collection Problems
+### Data Collection Concerns
 
@@
 This leaves out a big piece of the full picture on apartment quality, and may also be why some known wealthier wards are getting higher average scores as the low income housing there does not count.
-It also leaves out the effect of income inequality as lower income people are now more likely to live in such types of "apartments" than the wealthy [@Musterd2014].
-For apartments that did qualify, the grading is a scheduled event organized with the landlord ahead of time [@torontoeval].
-This means that scores presented may not be an entirely accurate measure of the general upkeep of a building as landlords who neglect their property can hurriedly make affairs before the inspector/enforcement officer arrives.
+It also leaves out the full effect of income inequality as lower income people are now more likely to live in such types of "apartments" than the wealthy [@Musterd2014].
+For apartments that do qualify for data collection, the grading is a scheduled event organized with the landlord ahead of time [@torontoeval].
+This means that scores presented may not be an entirely accurate measure of the general upkeep of a building as landlords who neglect their property can hurriedly make repairs and straighten out the apartment block before the inspector/enforcement officer arrives.
 It may thus be necessary for inspectors to come up semi-unannounced to get the true evaluation of the building.
@@
 It is important not to overlook those who are struggling to pay rent to access these improved amenities.
-Not all those who live in private areas or wealthier wards are rich, as wealthier wards still have poor areas so their own experiences may be neglected.
+Not all those who live in private areas or wealthier wards are rich as wealthier wards still have poor areas, so their own experiences may be neglected.
 
@@
 
-(should these be in next steps?)Private housing is now in itself getting harder to afford due to rising housing costs, forcing many so called middle class people into social housing, or lower rated private apartments.
-It would be interesting to see the effect of rising prices and the movement of people that would have been considered high income into apartments in lower income wards or social housing, on total maintenance scores and security, thus causing a hogher demand for those housing and tighter stretching available funds [@irwin].
+Private housing is now in itself getting harder to afford due to rising apartment costs, forcing many so called middle class people into social housing or lower rated private apartments.
+The effect of rising prices and the movement of people that would not have been considered low income into apartments in lower income wards or social housing apartments, causes a causing a higher demand for these apartments and thus causing a larger strain on already heavily strained available social housing funds [@irwin].
 
@@
 As average income was not provided for the data set, we had to collect external data which was set in 2016 and may not be accurate as it is not current to the modern day.
-
-While the N/As were dispersed randomly throughout the data set, removing them may have caused bias in the data set so our means, medians and analysis of Scores may have been different if the N/A values had been filled.
+While the N/As were dispersed randomly throughout the data set, removing them may have caused bias in the data set so our analysis of Scores, security and year built may have been different if the N/A values had been filled.
 Also, removing variables even if they were already present in the Score variable may have caused incorrect analysis as there may be some important information in the rejected variables that either brings a new perspective or aids one of the accepted variable in explaining the relationships between apartment type, ward, maintenance scores and security.
@@
 For next steps we could use the geo-location variable that was omitted for this report to see if the scores for apartments on boundaries between poorer or richer wards had the negative bias of the poor wards or positive bias of the rich.
-We could also use this variable for a geo-spatial legend in contrast to the graphical legend used in the report.
-We could also include apartment blocks that did not have all amenities in the comparison of income and scores to further observe evidence of inequality as apartments in poor areas tend to have complete absence of some amenities.
+We could also use this variable for a spatial comparison in contrast to the graphical comparison used in the report.
+We could also include apartment blocks that did not have all amenities in the comparison of income and scores to further observe evidence of inequality, as apartments in poor areas tend to have complete absence of some amenities.
 

### 2022-02-07 outputs/paper/paper.Rmd
Some final changes to the code.

@@
   kable_styling(latex_options = "HOLD_position")
+
 ```
@@
   theme_minimal() +
-  scale_color_manual(values = c("PRIVATE" = "brown", "SOCIAL HOUSING" = "black", "TCHC" = "darkorange")) + # custom colors
   labs(x = "Year Built",

### 2022-02-07 outputs/paper/paper.Rmd
Last changes on some grammatical errors.

@@
 We obtained the data set from the City of Toronto Open Data Portal [@citewebsite] with `opendatatoronto` [@citeopentoronto], and cleaned and analyzed the data set using `R` [@citeR], `dplyr`, [@citedplyr] and `tidyverse` [@citetidyverse] packages.
-Figures and tables were created with `ggplot2` [@citeggplot], `knitr` [@citeknit] and `kableExtra` [@citekable].
+Figures and tables were created with `ggplot2` [@citeggplot], `knitr` [@citeknit], `dplyr` [@citedplyr] and `kableExtra` [@citekable].
 We first created tables and graphs that showed the data.
@@
 
-The data for this report was gotten from the Apartment Building Evaluation [@citedata] of the Toronto Open Data Portal [@citewebsite] using the R package opendatatoronto [@citeopentoronto].
+The data for this report was gotten from the Apartment Building Evaluation [@citedata] of the Toronto Open Data Portal [@citewebsite] using the R package `opendatatoronto` [@citeopentoronto].
 This data is compiled by RentSafeTO and was last accessed on January 22nd, 2022.
 Rent Safe TO collects data on apartments in the Greater Toronto Area (GTA) that are 3 or more stories tall or had 10 more units [@torontoeval], and response is mandatory for qualifying apartments, i.e the population.
-RentSafeTO collects the data by setting up an appointment with the landlord of the apartment block for inspection, and take notes and photographs of their inspection to upload to the City’s mobile investigation application.
+RentSafeTO collects the data by setting up an appointment with the landlord of the apartment block for inspection, takes notes and photographs of their inspection to upload to the City’s mobile investigation application.
 Later, the total building evaluation score is calculated and made available to review and post on the Tenant Notification Board [@torontoeval].
@@
 We omitted most variables in the original set as their direct effects were not important to the analysis, and the total score variable encompassed all of their individual scores.
-we first removed _id and RSN as they were merely ID variables.
+We first removed _id and RSN as they were merely ID variables.
 We also excluded the numeric ward variable as the ward names variable, which contains the 25 Toronto wards, was easier to use for identification.
@@
 These relationships show a measure of income inequality in housing, and have possible implications for public policy.
-The data, (Table \@ref(tab:tabone)), (Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) using `kableextra` [@citekable], `knitr` [@citeknit] and `ggplot2` [@citeggplot] respectively are shown below.
+A glimpse of the data, (Table \@ref(tab:tabone)), using `kableextra` [@citekable], and `knitr` [@citeknit] are shown below.
 
@@
 
-(Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) show the spread and distribution of apartments scores and their security levels respectively.
+(Figure \@ref(fig:figone)) and (Figure \@ref(fig:figtwo)) using `ggplot2` [@citeggplot] show the spread and distribution of apartments scores and their security levels respectively.
 We see that the data for Scores seems to be unimodal and normally distributed.
-There is an outlier of a very low apartment score, but it is not enough to skew the data to the left.The median apartment score seems to be the same or a slightly above the mean apartment score, as there seems to be more apartments on the lower end than there are on the higher end.
+There is an outlier of a very low apartment score, but it is not enough to skew the data to the left.
+The median apartment score seems to be the same or a bit above the mean apartment score, as there seems to be more apartments on the lower end than there are on the higher end.
 This indicates that there are many below average apartments.
@@
   scale_y_continuous(labels = scales::percent)+ # getting the percentage of scores within each ward
-  scale_color_manual("YlOrRd")+
+  guides(fill=guide_legend(title = "Score Group"))+
   labs(x = "Ward",
@@
 
-The report has shown a disparity in priorities for apartment maintenance and upkeep between public and private apartments, as well as general housing inequality as those in social housing owned by the TCHC do not have the high apartment quality or security they deserve [@rights].
+The report has shown a disparity in priorities in apartment maintenance and upkeep between public and private apartments, as well as general housing inequality as those in social housing owned by the TCHC do not have the high apartment quality or security they deserve [@rights].
 Regarding these apartments that have less security, it is paradoxical as these people often live in more crime ridden areas that require the extra security [@Stretesky2004].
+Lower income people often do not have money to protect themselves like the wealthy, so they need to rely on publicly provided assistance.
 This brings into question if their own lives are considered, as well as the failings of public landlords and the TCHC to provide security akin to those of private apartments.
-Lower income people often do not have money to protect themselves like the wealthy, so they need to rely on publicly provided assistance.
 
 The data showed that Scarborough North, which has a 90% visible minority population with an average yearly income of 78,892 dollars as of 2016 compared to the Toronto average of 102,721 dollars [@torontoward], only has highly scored apartment buildings.
-If this is not due to a sampling error, it is hopeful that apartment scores for other wards with similar incomes increase in the future.
+If this is not due to a sampling problem, it is hopeful that apartment scores for other wards with similar incomes increase in the future.
 However low income residents have concerns that a high level of apartment security, or highly rated apartments in an otherwise lower income area can make the area more attractive to wealthier people.
@@
 There are some data collection problems with this data set.
-RentSafeTO collects data only for apartment blocks with 3 or more stories or ten or more units[@torontoeval], meaning that some buildings that are in dire need of maintenance may be left behind as the organization simply does not collect data for them, even though most people may see them as apartments (i.e a three story block of 9 flats).
+RentSafeTO collects data only for apartment blocks with 3 or more stories or ten or more units [@torontoeval], meaning that some buildings that are in dire need of maintenance may be left behind as the organization simply does not collect data for them, even though most people may see them as apartments (i.e a three story block of 9 flats).
 This leaves out a big piece of the full picture on apartment quality, and may also be why some known wealthier wards are getting higher average scores as the low income housing there does not count.
@@
 
-Private housing is now in itself getting harder to afford due to rising apartment costs, forcing many so called middle class people into social housing or lower rated private apartments.
-The effect of rising prices and the movement of people that would not have been considered low income into apartments in lower income wards or social housing apartments, causes a causing a higher demand for these apartments and thus causing a larger strain on already heavily strained available social housing funds [@irwin].
+Private housing is now itself getting harder to afford due to rising apartment costs, forcing many so called middle class people into social housing or lower rated private apartments.
+The effect of rising prices has led to the movement of people that would not have been considered low income into social housing apartments or apartments in lower income wards, causing a higher demand for these apartments and thus leading to a larger strain on already heavily strained available social housing funds [@irwin].
 
@@
 We could also use this variable for a spatial comparison in contrast to the graphical comparison used in the report.
-We could also include apartment blocks that did not have all amenities in the comparison of income and scores to further observe evidence of inequality, as apartments in poor areas tend to have complete absence of some amenities.
+We could then include apartment blocks that did not have all amenities in the comparison of income and scores to further observe evidence of inequality, as apartments in poor areas tend to have complete absence of some amenities.
 

### 2022-02-07 scripts/01-data_cleaning.R
Changed the date.

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Clean the Apartment Maintenance data downloaded from Open Data Toronto 
 # Author: Olaedo Okpareke
-# Data: 2 February 2022
+# Data: 6 February 2022
 # Contact: olaedo.okpareke@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 

### 2022-02-07 outputs/paper/paper.Rmd
Changed a few words.

@@
 
-## Data discussion
+# Discussion
 
-### Societal Impact of Scores
+## Societal Impact of Scores
 
@@
 
-### Bias and Ethical Concerns
+## Bias and Ethical Concerns
 
@@
 
-### Data Collection Concerns
+## Data Collection Concerns
 
@@
 
-### Income
+## Income
 
@@
 As average income was not provided for the data set, we had to collect external data which was set in 2016 and may not be accurate as it is not current to the modern day.
-While the N/As were dispersed randomly throughout the data set, removing them may have caused bias in the data set so our analysis of Scores, security and year built may have been different if the N/A values had been filled.
-Also, removing variables even if they were already present in the Score variable may have caused incorrect analysis as there may be some important information in the rejected variables that either brings a new perspective or aids one of the accepted variable in explaining the relationships between apartment type, ward, maintenance scores and security.
+While the missing values were dispersed randomly throughout the data set, removing them may have caused bias in the data set so our analysis of Scores, security and year built may have been different if the missing values had been filled.
+Also, removing variables even if they were already present in the Score variable may have caused incorrect analysis.
+This is because there may be some important information in the rejected variables that either brings a new perspective or aids one of the accepted variable in explaining the relationships between apartment type, ward, maintenance scores and security.
 

### 2022-02-07 outputs/paper/paper.Rmd
Made slight change to the title.

@@
 ---
-title: "Income Inequality is revealed in Torontonian Apartment Maintenance"
+title: "Income Inequality is Revealed in Torontonian Apartment Maintenance"
 author: "Olaedo Okpareke"

### 2022-02-07 outputs/paper/paper.Rmd
added a package citation.

@@
 
-We obtained the data set from the City of Toronto Open Data Portal [@citewebsite] with `opendatatoronto` [@citeopentoronto], and cleaned and analyzed the data set using `R` [@citeR], `dplyr`, [@citedplyr] and `tidyverse` [@citetidyverse] packages.
+We obtained the data set from the City of Toronto Open Data Portal [@citewebsite] with `opendatatoronto` [@citeopentoronto], and cleaned and analyzed the data set using `R` [@citeR], `dplyr`, [@citedplyr], `tidyr` [@citetidyr] and `tidyverse` [@citetidyverse] packages.
 Figures and tables were created with `ggplot2` [@citeggplot], `knitr` [@citeknit], `dplyr` [@citedplyr] and `kableExtra` [@citekable].
